---
title: "Connecticut elections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Connecticut elections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```
## What it covers

The Connecticut scraper returns election results from the CT Election Management
System (CTEMS) at `ctemspublic.tgstg.net`. It covers all elections available in
the CTEMS dropdown — general elections, primaries, runoffs, and special elections
— from 2016 to the present.

Results include an `election_level` column classifying each race as `"Federal"`,
`"State"`, or `"Local"`, and a `contest_outcome` column (`"Won"` / `"Lost"`) on
the statewide totals.

> **Note:** The CTEMS site is a JavaScript-rendered AngularJS SPA. All scraping
> uses a headless Chromium browser via Playwright. Scraping all 169 towns across
> all elections in a year typically takes 30–60 minutes per election year. Use
> `level = "state"` for a much faster statewide-only result.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | — | `"CT"`, `"connecticut"`, or `"Connecticut"` |
| `year_from` | `NULL` | Start year, inclusive; `NULL` = no lower bound |
| `year_to` | `NULL` | End year, inclusive; `NULL` = no upper bound |
| `level` | `"all"` | `"all"` — statewide + town; `"state"` — statewide only; `"town"` — town-level only |
| `max_workers` | `2L` | Parallel Chromium browsers for town scraping (one per county) |

---

## Examples

### Statewide + town results for a single year

```{r, purl=FALSE}
res <- scrape_elections(state = "CT", year_from = 2024, year_to = 2024)

# Statewide totals — one row per candidate per office
res$state %>%
  filter(election_level == "Federal") %>%
  select(election_name, office, candidate, party, votes, vote_pct, contest_outcome)

# Town-level results
res$town %>%
  filter(office == "President of the United States") %>%
  select(town, county, candidate, party, votes)
```

### Statewide totals only (faster — skips town scraping)

```{r, purl=FALSE}
state_df <- scrape_elections(
  state     = "CT",
  year_from = 2024,
  year_to   = 2024,
  level     = "state"
)

state_df %>%
  filter(contest_outcome == "Won") %>%
  select(election_name, election_level, office, candidate, party, votes, vote_pct)
```

### Town-level only

```{r, purl=FALSE}
town_df <- scrape_elections(
  state     = "CT",
  year_from = 2022,
  year_to   = 2022,
  level     = "town"
)

# Summarise votes by county for a specific race
town_df %>%
  filter(office == "United States Senator") %>%
  group_by(county, candidate, party) %>%
  summarise(votes = sum(votes, na.rm = TRUE), .groups = "drop") %>%
  arrange(county, desc(votes))
```

### Multi-year scrape with more parallel workers

```{r, purl=FALSE}
res <- scrape_elections(
  state             = "CT",
  year_from         = 2020,
  year_to           = 2024,
  level             = "all",
  max_workers = 4L
)
```

### Filter by election level

```{r, purl=FALSE}
res <- scrape_elections(state = "CT", year_from = 2023, year_to = 2023)

# Local races only (municipal elections)
res$state %>%
  filter(election_level == "Local") %>%
  select(election_name, office, candidate, party, votes, contest_outcome)
```

---

## Output columns

### `$state` (when `level = "all"` or `"state"`)

| Column | Description |
|---|---|
| `election_name` | Human-readable election name (e.g. `"11/05/2024 -- November General Election"`) |
| `election_year` | Calendar year |
| `election_date` | ISO date string when parseable; `NA` otherwise |
| `election_level` | `"Federal"`, `"State"`, or `"Local"` |
| `office` | Office name (includes district when present) |
| `candidate` | Candidate name |
| `party` | Party name (e.g. `"Democratic"`, `"Republican"`) |
| `votes` | Total votes (integer) |
| `vote_pct` | Percentage of votes in contest (recomputed from town totals for State/Local) |
| `winner` | `TRUE` if the candidate won the statewide contest |

### `$town` (when `level = "all"` or `"town"`)

| Column | Description |
|---|---|
| `election_name` | Human-readable election name |
| `election_year` | Calendar year |
| `election_date` | Date of the election |
| `county` | County name |
| `town` | Town name |
| `office_level` | `"Federal"`, `"State"`, or `"Local"` |
| `office` | Office name |
| `candidate` | Candidate name |
| `party` | Party name |
| `votes` | Votes in this town (integer) |
| `vote_pct` | Percentage as reported by CTEMS for this town |
| `town_winner` | `TRUE` if the candidate won within this town |

---

## Performance notes

- **`level = "state"` is much faster.** Town scraping opens one Chromium browser
  per county and iterates all 169 towns. A single election year with full town
  scraping takes roughly 30–60 minutes. Use `level = "state"` when you only need
  statewide totals.
- **Multiple elections per year.** CTEMS often has several elections in a single
  year (primary, general, special elections). All are scraped and combined in the
  output.
- **`max_workers`.** CT has 8 counties. The default of 2 parallel browsers
  means 4 batches. Increasing to 4 roughly halves town-scraping time. Keep at or
  below 4 to avoid memory pressure.
