---
title: "ElectionStats states"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ElectionStats states}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```
## What it covers

The ElectionStats scraper pulls candidate-level and county-level election
results from state-hosted ElectionStats sites.

| State | Backend | Start year | End year |
|---|---|---|---|
| Vermont | Classic (requests) | 1789 | 2024 |
| Virginia | Classic (requests) | 1789 | 2025 |
| Colorado | Classic (requests) | 1902 | 2024 |
| Massachusetts | Classic (requests) | 1970 | 2026 |
| New Hampshire | Classic (requests) | 1970 | 2024 |
| Idaho | Classic (requests) | 1990 | 2024 |
| New York | v2 (Playwright) | 1994 | 2024 |
| New Mexico | v2 (Playwright) | 2000 | 2024 |
| South Carolina | v2 (Playwright) | 2008 | 2025 |

Classic states use simple HTTP requests and are fast. Playwright states
(NY, NM, SC) launch a headless Chromium browser to render JavaScript —
they are slower but fully automated.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | required | State name, snake\_case (e.g. `"virginia"`) |
| `year_from` | `NULL` (earliest) | Start year, inclusive |
| `year_to` | `NULL` (current year) | End year, inclusive |
| `level` | `"all"` | What to return — see below |
| `parallel` | `TRUE` | Parallel county scraping (classic states only) |

### The `level` parameter

| Value | Returns |
|---|---|
| `"all"` (default) | Named list with `$state`, `$county`, and (where available) `$precinct` data frames |
| `"state"` | Candidate-level results data frame |
| `"county"` | County vote breakdown data frame |
| `"precinct"` | Precinct-level vote breakdown data frame (CO, MA, ID, SC, NM, VA) |

---

## Examples

### Candidate-level results for a single year

```{r, purl=FALSE}
state_df <- scrape_elections(
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  level     = "state"
)

state_df %>%
  filter(winner == TRUE) %>%
  arrange(desc(votes)) %>%
  select(office, district, candidate, party, votes)
```

### County-level vote counts

```{r, purl=FALSE}
county_df <- scrape_elections(
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  level     = "county"
)

county_df %>%
  arrange(desc(votes)) %>%
  select(county_or_city, candidate, votes)
```

### Both levels at once (default)

`level = "all"` returns a named list with `$state`, `$county`, and (for precinct-capable states) `$precinct` elements:

```{r, purl=FALSE}
res <- scrape_elections(
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023
)

res$state   # candidate-level data frame
res$county  # county breakdown data frame
```

### Multi-year range

```{r, purl=FALSE}
ma_results <- scrape_elections(
  state     = "massachusetts",
  year_from = 2018,
  year_to   = 2022,
  level     = "state"
)

ma_results %>%
  filter(election_type == "General") %>%
  select(election_year, office, district, candidate, party, votes)
```

### Playwright state (South Carolina)

```{r, purl=FALSE}
sc_results <- scrape_elections(
  state     = "south_carolina",
  year_from = 2024,
  year_to   = 2024,
  level     = "state"
)

sc_results %>%
  filter(winner == TRUE) %>%
  select(office, district, candidate, party, vote_pct)
```

### Precinct-level results

Available for CO, MA, ID (classic/requests-based) and SC, NM, VA (v2 — fetched
via CSV download API).  Not available for NH, VT, or NY.

```{r, purl=FALSE}
# Classic state
precinct_df <- scrape_elections(
  state     = "colorado",
  year_from = 2022,
  year_to   = 2022,
  level     = "precinct"
)

precinct_df %>%
  arrange(county, precinct) %>%
  select(county, precinct, candidate, votes)
```

```{r, purl=FALSE}
# v2 state — same interface, data fetched via CSV API
precinct_df <- scrape_elections(
  state     = "new_mexico",
  year_from = 2024,
  year_to   = 2024,
  level     = "precinct"
)
```

Or retrieve all levels at once and access `$precinct`:

```{r, purl=FALSE}
res <- scrape_elections(
  state     = "south_carolina",
  year_from = 2024,
  year_to   = 2024
)

res$precinct  # precinct-level data frame (NULL for NH, VT, NY)
```

### Disable parallel scraping

```{r, purl=FALSE}
res_seq <- scrape_elections(
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  parallel  = FALSE
)
```

---

## Columns returned

### `level = "state"` (candidate-level)

- `state`, `election_year`, `election_id`, `election_type`
- `office_level`, `office`, `district`
- `candidate_id`, `candidate`, `party`
- `votes`, `vote_pct`, `winner`
- `url`

### `level = "county"`

- `state`, `election_year`, `election_id`, `election_type`
- `office_level`, `office`, `district`, `county_or_city`
- `candidate_id`, `candidate`, `party`
- `votes`, `vote_pct`
- `county_winner` — `TRUE` for the top vote-getter within each county/city
- `url`

### `level = "precinct"` (CO, MA, ID, SC, NM, VA)

- `state`, `election_year`, `election_id`, `election_type`
- `office_level`, `office`, `district`, `county`, `precinct`
- `candidate_id`, `candidate`, `party`
- `votes`, `vote_pct`
- `precinct_winner` — `TRUE` for the top vote-getter within each precinct
- `url`
