---
title: "ElectionStats states"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ElectionStats states}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

```{r setup}
library(DownBallotR)
library(dplyr)

downballot_use_python()
```

## What it covers

The ElectionStats scraper pulls candidate-level and county-level election
results from state-hosted ElectionStats sites.

| State | Backend | Start year | End year |
|---|---|---|---|
| Vermont | Classic (requests) | 1789 | 2024 |
| Virginia | Classic (requests) | 1789 | 2025 |
| Colorado | Classic (requests) | 1902 | 2024 |
| Massachusetts | Classic (requests) | 1970 | 2026 |
| New Hampshire | Classic (requests) | 1970 | 2024 |
| New York | v2 (Playwright) | 1994 | 2024 |
| New Mexico | v2 (Playwright) | 2000 | 2024 |
| South Carolina | v2 (Playwright) | 2008 | 2025 |

Classic states use simple HTTP requests and are fast. Playwright states
(NY, NM, SC) launch a headless Chromium browser to render JavaScript —
they are slower but fully automated.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | required | State name, snake\_case (e.g. `"virginia"`) |
| `year_from` | `NULL` (earliest) | Start year, inclusive |
| `year_to` | `NULL` (current year) | End year, inclusive |
| `level` | `"all"` | What to return — see below |
| `parallel` | `TRUE` | Parallel county scraping (classic states only) |

### The `level` parameter

| Value | Returns |
|---|---|
| `"all"` (default) | Named list with `$state` and `$county` data frames |
| `"state"` | Candidate-level results data frame |
| `"county"` | County vote breakdown data frame |
| `"joined"` | County rows merged with candidate metadata |

---

## Examples

### Candidate-level results for a single year

```{r}
state_df <- scrape_elections(
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  level     = "state"
)

state_df %>%
  filter(contest_outcome == "Winner") %>%
  arrange(desc(total_vote_count)) %>%
  select(office, district, candidate, party, total_vote_count)
```

### County-level vote counts

```{r}
county_df <- scrape_elections(
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  level     = "county"
)

county_df %>%
  arrange(desc(votes)) %>%
  select(county_or_city, candidate_name, votes)
```

### Both levels at once (default)

`level = "all"` returns a named list with `$state` and `$county` elements:

```{r}
res <- scrape_elections(
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023
)

res$state   # candidate-level data frame
res$county  # county breakdown data frame
```

### Multi-year range

```{r}
ma_results <- scrape_elections(
  state     = "massachusetts",
  year_from = 2018,
  year_to   = 2022,
  level     = "state"
)

ma_results %>%
  filter(stage == "General") %>%
  select(year, office, district, candidate, party, total_vote_count)
```

### Playwright state (South Carolina)

```{r}
sc_results <- scrape_elections(
  state     = "south_carolina",
  year_from = 2024,
  year_to   = 2024,
  level     = "state"
)

sc_results %>%
  filter(contest_outcome == "Winner") %>%
  select(office, district, candidate, party, vote_percentage)
```

### Joined data (county + candidate metadata)

```{r}
joined <- scrape_elections(
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  level     = "joined"
)

joined %>%
  select(county_or_city, office, candidate, party, votes)
```

### Disable parallel scraping

```{r}
res_seq <- scrape_elections(
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  level     = "joined",
  parallel  = FALSE
)
```

---

## Columns returned

### `level = "state"` (candidate-level)

- `state`, `year`, `election_id`, `candidate_id`
- `office`, `district`, `stage` (General, Primary, etc.)
- `candidate`, `party`
- `total_vote_count`, `vote_percentage`, `contest_outcome`
- `detail_url`

### `level = "county"`

- `state`, `year`, `election_id`, `candidate_id`
- `county_or_city`, `candidate_name`, `votes`
