---
title: "Indiana elections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Indiana elections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```
## What it covers

The Indiana scraper returns General Election results from the Indiana Election
Division at `enr.indianavoters.in.gov`. It covers General Elections only
(not primaries or special elections) from 2019 to the present.

Results are fetched from the FirstTuesday archive JSON API and include
statewide candidate totals and county-level breakdowns for all Federal,
State, and Local office categories.

> **Note:** The Indiana site is a JSON API (no browser required). Scraping
> is fast relative to Playwright-based states. County-level data covers all
> 92 Indiana counties.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | — | `"IN"`, `"indiana"`, or `"Indiana"` |
| `year_from` | `NULL` | Start year, inclusive; `NULL` = no lower bound (clamped to 2019) |
| `year_to` | `NULL` | End year, inclusive; `NULL` = current calendar year |
| `level` | `"all"` | `"all"` — statewide + county; `"state"` — statewide only; `"county"` — county-level only |

---

## Examples

### Statewide + county results for a single year

```{r, purl=FALSE}
res <- scrape_elections(state = "IN", year_from = 2024, year_to = 2024)

# Statewide candidate totals
res$state %>%
  filter(office_level == "Federal") %>%
  select(election_year, office_title, candidate, party, votes, vote_pct, winner)

# County-level breakdown
res$county %>%
  filter(office_title == "President of the United States") %>%
  select(county_name, candidate, party, votes, vote_pct) %>%
  arrange(county_name, desc(votes))
```

### Statewide totals only (faster — skips county scraping)

```{r, purl=FALSE}
state_df <- scrape_elections(
  state     = "IN",
  year_from = 2024,
  year_to   = 2024,
  level     = "state"
)

state_df %>%
  select(election_year, office_level, office_title, candidate, party, votes, vote_pct, winner)
```

### Multi-year scrape

```{r, purl=FALSE}
res <- scrape_elections(
  state     = "IN",
  year_from = 2020,
  year_to   = 2024,
  level     = "state"
)

res %>%
  filter(office_title == "President of the United States") %>%
  select(election_year, candidate, party, votes, vote_pct, winner)
```

### Filter to a specific county

```{r, purl=FALSE}
res <- scrape_elections(state = "IN", year_from = 2024, year_to = 2024)

res$county %>%
  filter(county_name == "Marion") %>%
  select(office_title, candidate, party, votes, vote_pct) %>%
  arrange(office_title, desc(votes))
```

### Filter by office level

```{r, purl=FALSE}
res <- scrape_elections(state = "IN", year_from = 2024, year_to = 2024)

# Federal races only
res$state %>%
  filter(office_level == "Federal") %>%
  select(office_title, candidate, party, votes, vote_pct, winner)

# State races only
res$state %>%
  filter(office_level == "State") %>%
  select(office_title, candidate, party, votes, vote_pct, winner)
```

---

## Output columns

### `$state` (when `level = "all"` or `"state"`)

| Column | Description |
|---|---|
| `state` | State abbreviation (`"IN"`) |
| `election_year` | Calendar year of the election |
| `election_date` | Date of the election |
| `election_type` | Always `"General"` |
| `office_level` | `"Federal"`, `"State"`, or `"Local"` |
| `office_category` | Broad category name (e.g. `"County Offices"`) |
| `office_title` | Specific office name (e.g. `"President of the United States"`) |
| `office_id` | Internal office identifier |
| `candidate` | Candidate name |
| `candidate_name_on_ballot` | Name as printed on the ballot |
| `party` | Full party name (e.g. `"Republican"`, `"Democrat"`) |
| `votes` | Statewide vote total |
| `vote_pct` | Percentage of votes within the contest |
| `winner` | `TRUE` if the candidate won |
| `num_seats` | Number of seats available in the contest |
| `precincts_reporting` | Total precincts reporting statewide |
| `total_precincts` | Total precincts statewide |

### `$county` (when `level = "all"` or `"county"`)

Same columns as `$state`, except `precincts_reporting` and `total_precincts`
are absent, `winner` is replaced by `county_winner`, and two county-specific
columns are added:

| Column | Description |
|---|---|
| `county_name` | County name |
| `county_fips` | County FIPS code |
| `county_winner` | `TRUE` if the candidate won within this county |

---

## Performance notes

- **Fast scraper.** Indiana uses a JSON API — no headless browser is required.
  A full statewide + county scrape for one year typically completes in under a
  minute.
- **General Elections only.** The scraper covers General Elections only.
  Primary and special election results are not available through this source.
- **Local races.** County-level data includes Local office races (county
  treasurer, clerk, etc.) in addition to Federal and State races.
