---
title: "Municipal and mayoral elections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Municipal and mayoral elections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

```{r setup}
library(DownBallotR)
library(dplyr)

downballot_use_python()
```

## What it covers

The Ballotpedia municipal elections scraper collects city, county, and mayoral
election data from two Ballotpedia index pages. Use
`office = "municipal_elections"` to activate it.

Two index pages are available via the `race_type` argument:

| `race_type` | Ballotpedia page | Coverage |
|---|---|---|
| `"all"` (default) | `United_States_municipal_elections,_{year}` | City council, county, and mayoral races; 2014–present |
| `"mayoral"` | `United_States_mayoral_elections,_{year}` | Mayoral races only; 2020–present |

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `office` | — | Must be `"municipal_elections"` |
| `state` | `NULL` | State name (title-case, e.g. `"Texas"`), or `NULL` for all states |
| `year` | `NULL` | Single election year; use with `mode = "links"` or `"results"` |
| `race_type` | `"all"` | Which index page to use — see above |
| `mode` | `"links"` | What to return — see below |
| `start_year` | `2014` | Earliest year for a multi-year scrape when `year` is `NULL` |
| `end_year` | `NULL` | Latest year (default: current calendar year) |

### The `mode` parameter

| Value | Returns | Speed |
|---|---|---|
| `"links"` (default) | One row per election sub-URL with location metadata; no vote data | Fast — 1–2 requests per year |
| `"results"` | Full candidate data with vote counts and winner flags | Slower — 1 extra request per election page |

---

## Examples

### Index links for one year and state (fast)

```{r}
links <- scrape_elections(
  office = "municipal_elections",
  year   = 2022,
  state  = "Texas"
)

links %>%
  select(location, state, location_type, url) %>%
  arrange(location_type, location)
```

### All states, one year

```{r}
links_all <- scrape_elections(
  office = "municipal_elections",
  year   = 2022
)

# Breakdown by location type
links_all %>%
  count(location_type, sort = TRUE)
```

### Mayoral index links only

Use `race_type = "mayoral"` to pull only from the dedicated mayoral
index page (available 2020–present):

```{r}
mayoral_links <- scrape_elections(
  office    = "municipal_elections",
  year      = 2024,
  race_type = "mayoral"
)

mayoral_links %>%
  select(location, state, url) %>%
  arrange(state, location)
```

### Full candidate results for one year + state

`mode = "results"` follows each sub-URL — slower but returns candidate
names, vote counts, and winner flags:

```{r}
results <- scrape_elections(
  office = "municipal_elections",
  year   = 2022,
  state  = "Texas",
  mode   = "results"
)

results %>%
  filter(is_winner) %>%
  select(location, office, election_type, candidate, party, pct, votes) %>%
  arrange(location, office)
```

### Mayoral results only — one year, all states

```{r}
scrape_elections(
  office    = "municipal_elections",
  year      = 2022,
  race_type = "mayoral",
  mode      = "results"
) %>%
  filter(is_winner, office == "Mayor") %>%
  select(location, state, candidate, party, pct, votes) %>%
  arrange(state, location)
```

### Mayoral results for a single state

```{r}
scrape_elections(
  office    = "municipal_elections",
  year      = 2024,
  state     = "California",
  race_type = "mayoral",
  mode      = "results"
) %>%
  filter(office == "Mayor") %>%
  select(location, candidate, party, is_winner, pct, votes) %>%
  arrange(location)
```

### Multi-year mayoral results

```{r}
scrape_elections(
  office     = "municipal_elections",
  race_type  = "mayoral",
  mode       = "results",
  start_year = 2020,
  end_year   = 2022
) %>%
  filter(is_winner, office == "Mayor") %>%
  select(year, location, state, candidate, party, pct, votes) %>%
  arrange(year, state, location)
```

### Multi-year index links

```{r}
scrape_elections(
  office     = "municipal_elections",
  start_year = 2020,
  end_year   = 2022,
  race_type  = "all"
) %>%
  group_by(year, location_type) %>%
  summarise(n_elections = n(), .groups = "drop")
```

---

## Columns returned

### `mode = "links"`

- `year`, `location` (city or county name), `state`
- `location_type` — `"mayoral"`, `"city"`, `"county"`, or `"other"`
- `url` — full Ballotpedia URL of the individual election page

### `mode = "results"`

All link columns plus:

- `election_url` — URL of the individual election page
- `office` — e.g. `"Mayor"`, `"City Council"`, `"Public Defender"`
- `election_type` — `"General"`, `"Primary"`, `"Primary Runoff"`,
  `"General (RCV)"`, or `"Other"`
- `candidate`, `ballotpedia_url`, `party`
- `is_winner`, `is_incumbent`
- `pct` — vote percentage (e.g. `"62.3%"`) or `""`
- `votes` — vote count (e.g. `"12,345"`) or `""`
