---
title: "North Carolina elections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{North Carolina elections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```
## What it covers

The NC scraper returns precinct-level local election results from the
NC State Board of Elections, filtered by year range.

> **Note:** Only the 2025 election file layout is currently supported.
> Elections before 2025 use an older format that has not yet been
> implemented; they are automatically skipped with an informational message.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | `"NC"` (any NC alias) | `"NC"`, `"north_carolina"`, or `"north carolina"` |
| `year_from` | `NULL` | Start year, inclusive; `NULL` = no lower bound |
| `year_to` | `NULL` | End year, inclusive; `NULL` = no upper bound |

---

## Examples

### Precinct + county + state results for a single year

```{r, purl=FALSE}
res <- scrape_elections(state = "NC", year_from = 2025, year_to = 2025)

# Precinct-level results
res$precinct %>%
  filter(county == "Wake") %>%
  select(election_date, contest_name, precinct, candidate, party, votes) %>%
  arrange(contest_name, desc(votes))

# County-level aggregates
res$county %>%
  filter(office_level == "State") %>%
  select(county, election_date, contest_name, candidate, party, votes) %>%
  arrange(contest_name, county, desc(votes))

# Statewide totals
res$state %>%
  filter(winner == TRUE) %>%
  select(election_date, contest_name, candidate, party, votes, vote_pct)
```

### Statewide totals only (faster — skips precinct/county detail)

```{r, purl=FALSE}
state_df <- scrape_elections(state = "NC", year_from = 2025, year_to = 2025,
                              level = "state")

state_df %>%
  select(election_date, contest_name, candidate, party, votes, vote_pct, winner)
```

### Filter to a specific county

```{r, purl=FALSE}
res <- scrape_elections(state = "NC", year_from = 2025, year_to = 2025)

res$county %>%
  filter(county == "Wake") %>%
  select(election_date, contest_name, candidate, party, votes) %>%
  arrange(contest_name, desc(votes))
```

---

## Columns returned

### `$precinct` (when `level = "all"` or `"precinct"`)

| Column | Description |
|---|---|
| `state` | Always `"NC"` |
| `election_date` | Date of the election |
| `election_year` | Calendar year of the election |
| `election_type` | Election type (e.g. `"General"`, `"Primary"`) |
| `county` | County name |
| `precinct` | Precinct name or identifier |
| `contest_name` | Full contest name as reported by NCSBE |
| `contest_type` | Contest type code |
| `contest_group_id` | Internal contest group identifier |
| `jurisdiction` | Jurisdiction extracted from contest name |
| `office_level` | `"Federal"`, `"State"`, or `"Local"` |
| `district` | District identifier (when applicable) |
| `candidate` | Candidate name |
| `party` | Party abbreviation |
| `vote_for` | Number of seats to fill |
| `election_day` | Election day votes |
| `early_voting` | Early voting votes |
| `absentee_by_mail` | Absentee by mail votes |
| `absentee_or_early` | Combined absentee/early votes (older files) |
| `provisional` | Provisional votes |
| `votes` | Total votes |
| `winner_status` | Raw winner status string from NCSBE source |
| `runoff_status` | Runoff indicator from source |
| `recount_status` | Recount indicator from source |
| `real_precinct` | Whether the row represents a real precinct |
| `precinct_abbrv` | Precinct abbreviation |
| `ftp_date` | File date from NCSBE |

### `$county` (when `level = "all"` or `"county"`)

| Column | Description |
|---|---|
| `state` | Always `"NC"` |
| `county` | County name |
| `election_date` | Date of the election |
| `contest_name` | Full contest name |
| `jurisdiction` | Jurisdiction extracted from contest name |
| `office_level` | `"Federal"`, `"State"`, or `"Local"` |
| `district` | District identifier (when applicable) |
| `candidate` | Candidate name |
| `party` | Party abbreviation |
| `election_day` | Election day votes aggregated to county |
| `early_voting` | Early voting votes aggregated to county |
| `absentee_by_mail` | Absentee by mail votes aggregated to county |
| `absentee_or_early` | Combined absentee/early votes (older files) |
| `provisional` | Provisional votes aggregated to county |
| `votes` | Total votes aggregated to county |

### `$state` (when `level = "all"` or `"state"`)

Same columns as `$county`, minus `county`, plus:

| Column | Description |
|---|---|
| `vote_pct` | Percentage of votes within the contest |
| `winner` | `TRUE` for the highest-voted candidate(s) per contest |
