---
title: "North Carolina elections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{North Carolina elections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```
## What it covers

The NC scraper returns precinct-level local election results from the
NC State Board of Elections, filtered by year range.

> **Note:** Only the 2025 election file layout is currently supported.
> Elections before 2025 use an older format that has not yet been
> implemented; they are automatically skipped with an informational message.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | `"NC"` (any NC alias) | `"NC"`, `"north_carolina"`, or `"north carolina"` |
| `year_from` | `NULL` | Start year, inclusive; `NULL` = no lower bound |
| `year_to` | `NULL` | End year, inclusive; `NULL` = no upper bound |

---

## Examples

### All currently supported NC results

```{r, purl=FALSE}
nc_all <- scrape_elections(state = "NC")

nc_all %>%
  select(year, county, office, candidate, party, votes) %>%
  arrange(county, office)
```

### Single year

```{r, purl=FALSE}
nc_2025 <- scrape_elections(
  state     = "NC",
  year_from = 2025,
  year_to   = 2025
)

nc_2025 %>%
  filter(contest_outcome == "Winner") %>%
  select(county, office, candidate, party, votes)
```

### Filter to a specific county

```{r, purl=FALSE}
scrape_elections(state = "NC") %>%
  filter(county == "Wake") %>%
  select(year, office, candidate, party, votes) %>%
  arrange(office, desc(votes))
```

### Multi-year range

```{r, purl=FALSE}
nc_range <- scrape_elections(
  state     = "NC",
  year_from = 2025,
  year_to   = 2026
)

nc_range %>%
  group_by(year, county) %>%
  summarise(races = n_distinct(election_id), .groups = "drop")
```

---

## Columns returned

- `state`, `year`, `election_id`
- `county`, `precinct`
- `office`, `district`, `stage`
- `candidate`, `party`
- `votes`, `contest_outcome`
