---
title: "School district elections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{School district elections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

```{r setup}
library(DownBallotR)
library(dplyr)

downballot_use_python()
```

## What it covers

The school district scraper collects school board election data from
Ballotpedia's annual school board election pages (2013 to present,
all US states). Use `office = "school_district"` to activate it.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `office` | — | Must be `"school_district"` |
| `state` | `NULL` | State name (e.g. `"Alabama"`), or `NULL` for all states |
| `year` | `NULL` | Single election year. Required for `mode = "results"` or `"joined"`. |
| `mode` | `"districts"` | What to return — see below |
| `start_year` | `2013` | Earliest year for a multi-year district scrape |
| `end_year` | `NULL` | Latest year (default: current calendar year) |

### The `mode` parameter

| Value | Returns | Speed |
|---|---|---|
| `"districts"` (default) | District-level metadata — one request per year page | Fast |
| `"results"` | Candidate names, votes, percentages per race | Slower — one extra request per district |
| `"joined"` | Districts + candidates in a single merged data frame | Slower |

---

## Examples

### District metadata for all states

```{r}
districts_2024 <- scrape_elections(
  office = "school_district",
  year   = 2024
)

districts_2024 %>%
  filter(as.integer(seats_up) > 3) %>%
  select(state, district, seats_up, total_board_seats, enrollment) %>%
  arrange(desc(as.integer(seats_up)))
```

### Filter to one state

```{r}
alabama_2024 <- scrape_elections(
  state  = "Alabama",
  office = "school_district",
  year   = 2024
)

alabama_2024 %>%
  select(district, general_election, seats_up, total_board_seats, enrollment)
```

### Competitive districts (many seats up)

```{r}
scrape_elections(
  state  = "Texas",
  office = "school_district",
  year   = 2024
) %>%
  filter(as.integer(seats_up) > 1) %>%
  select(district, general_election, seats_up, total_board_seats, enrollment) %>%
  arrange(desc(as.integer(seats_up)))
```

### Full candidate results for one year + state

`mode = "results"` follows each district link — slower but returns
candidate names, vote counts, and winner flags:

```{r}
results <- scrape_elections(
  state  = "New Jersey",
  office = "school_district",
  year   = 2024,
  mode   = "results"
)

# Winners only
results %>%
  filter(is_winner) %>%
  select(district, race, candidate, party, pct, votes) %>%
  arrange(district, race)
```

### Districts + candidates joined

```{r}
joined <- scrape_elections(
  state  = "Texas",
  office = "school_district",
  year   = 2024,
  mode   = "joined"
)

joined %>%
  filter(is_winner) %>%
  select(district, enrollment, candidate, party, pct, votes)
```

### Multi-year district metadata

Use `start_year` / `end_year` when `year` is omitted:

```{r}
multi_year <- scrape_elections(
  state      = "Ohio",
  office     = "school_district",
  start_year = 2020,
  end_year   = 2024
)

multi_year %>%
  group_by(year) %>%
  summarise(n_districts = n(), .groups = "drop")
```

---

## Columns returned

### `mode = "districts"`

- `year`, `state`, `district`, `district_url`
- `primary`, `primary_runoff`, `general_election`, `general_runoff`
- `term_length`, `seats_up`, `total_board_seats`, `enrollment`

### `mode = "results"` / `"joined"`

All district columns plus:

- `race`, `election_type` (General / Primary / Primary Runoff / Other)
- `candidate`, `candidate_url`, `party`
- `is_winner`, `is_incumbent`
- `pct`, `votes`
