---
title: "Scraping election data with DownBallotR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scraping election data with DownBallotR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Prerequisites

Before scraping data you need a working Python environment. If you have not
done so yet, follow the steps in the
[Python setup vignette](python-setup.html):

```{r}
# One-time setup (downloads ~100-200 MB the first time)
downballot_install_python()
```

At the start of each R session:

```{r}
library(DownBallotR)
library(dplyr)

downballot_use_python()
```

---

## The single entry point: `scrape_elections()`

All data retrieval flows through one function:

```r
scrape_elections(state, office = "general", ...)
```

The scraper is selected **automatically** based on `state` and `office` —
you never need to specify a backend by name:

| Priority | Condition | Scraper used |
|---|---|---|
| 1 | `office = "school_district"` | Ballotpedia school board scraper (all US states) |
| 2 | `office = "state_elections"` | Ballotpedia state elections (2024–present) |
| 3 | `office = "municipal_elections"` | Ballotpedia municipal + mayoral elections (2014–present) |
| 4 | `state` is North Carolina | NC State Board of Elections scraper |
| 5 | All other states | ElectionStats multi-state scraper |

---

## Data availability

Use `db_available_years()` to check available year ranges programmatically:

```{r}
# All sources
db_available_years()

# Filter to one office type
db_available_years(office = "school_district")
db_available_years(office = "state_elections")
db_available_years(office = "municipal_elections")

# Filter to one state
db_available_years(state = "virginia")
```

| Source | State | Start year | End year |
|---|---|---|---|
| ElectionStats | Vermont | 1789 | 2024 |
| ElectionStats | Virginia | 1789 | 2025 |
| ElectionStats | Colorado | 1902 | 2024 |
| ElectionStats | Massachusetts | 1970 | 2026 |
| ElectionStats | New Hampshire | 1970 | 2024 |
| ElectionStats | New York | 1994 | 2024 |
| ElectionStats | New Mexico | 2000 | 2024 |
| ElectionStats | South Carolina | 2008 | 2025 |
| NC State Board | North Carolina | 2025 | 2025 |
| Ballotpedia (school boards) | All US states | 2013 | present |
| Ballotpedia (state elections) | All US states\* | 2024 | present |
| Ballotpedia (municipal elections) | All US states | 2014 (`"all"`) / 2020 (`"mayoral"`) | present |

To see which states are supported by the ElectionStats scraper:

```{r}
db_list_states("election_stats")
#> [1] "colorado"       "massachusetts"  "new_hampshire"  "new_mexico"
#> [5] "new_york"       "south_carolina" "vermont"        "virginia"
```

---

## Choosing the right scraper

| Goal | Call |
|---|---|
| Candidate + vote totals for VA, MA, CO, NH, VT | `scrape_elections(state = "virginia", ...)` |
| Candidate + vote totals for SC, NM, NY (Playwright) | `scrape_elections(state = "south_carolina", ...)` |
| NC precinct-level local results | `scrape_elections(state = "NC", year_from = 2025, year_to = 2025)` |
| School board district schedules, any state | `scrape_elections(office = "school_district", mode = "districts", ...)` |
| School board candidates + results | `scrape_elections(office = "school_district", mode = "results", ...)` |
| All candidates for a state (federal + state + local), 2024+ | `scrape_elections(state = "Maine", office = "state_elections", year = 2024)` |
| Federal candidates only, with vote counts | `scrape_elections(state = "Maine", office = "state_elections", year = 2024, election_level = "federal", mode = "results")` |
| Discover municipal/mayoral election URLs for a year | `scrape_elections(office = "municipal_elections", year = 2022)` |
| Full candidate results for municipal elections | `scrape_elections(office = "municipal_elections", year = 2022, mode = "results")` |
| Multi-year mayoral results, one state | `scrape_elections(office = "municipal_elections", state = "Texas", race_type = "mayoral", mode = "results", start_year = 2020, end_year = 2024)` |

---

## Detailed vignettes

Each scraper has its own vignette with full argument documentation,
worked examples, and column descriptions:

- [ElectionStats states](election-stats.html) —
  VA, MA, CO, NH, SC, NM, NY, VT; candidate + county results
- [North Carolina](north-carolina.html) —
  NC State Board of Elections; precinct-level results
- [School district elections](school-district-elections.html) —
  Ballotpedia school board elections, all US states, 2013–present
- [State elections](state-elections.html) —
  Ballotpedia federal/state/local candidate listings, 2024–present
- [Municipal elections](municipal-elections.html) —
  Ballotpedia city, county, and mayoral races, 2014–present

---

## Performance tips

- **Start small** — test with a single year and state before requesting
  large date ranges.
- **Playwright states are slower** — South Carolina, New Mexico, and New York
  launch a headless browser for each scrape. Expect several seconds per year.
- **Ballotpedia `mode = "results"`** makes one extra HTTP request per
  district, contest, or election page. Use `mode = "districts"` /
  `mode = "listings"` / `mode = "links"` for a fast overview without
  vote counts.
- **Parallel scraping is on by default** for classic (requests-based)
  ElectionStats states. Pass `parallel = FALSE` to disable.
- **Be polite** — built-in delays are intentional. Do not reduce or
  disable them; excessive requests may result in temporary IP blocks.
