---
title: "Scraping election data with DownBallotR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scraping election data with DownBallotR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Prerequisites

Before scraping data you need a working Python environment. If you have not
done so yet, follow the steps in the
[Python setup vignette](python-setup.html):

```{r}
# One-time setup (downloads ~100-200 MB the first time)
downballot_install_python()
```

At the start of each R session:

```{r}
library(DownBallotR)
downballot_use_python()
```

---

## The single entry point: `scrape_elections()`

All data retrieval flows through one function:

```r
scrape_elections(source, ...)
```

The `source` argument selects the backend scraper. Every other argument
is source-specific and documented in the sections below.

To see which sources and states are available:

```{r}
db_list_sources()
#> [1] "ballotpedia"    "election_stats" "nc_results"

db_list_states("election_stats")
#> [1] "colorado"       "massachusetts"  "new_hampshire"  "new_mexico"
#> [5] "new_york"       "south_carolina" "virginia"
```

---

## Source 1 — ElectionStats (`source = "election_stats"`)

### What it covers

The ElectionStats scraper pulls candidate-level and county-level election
results from state-hosted ElectionStats sites. Supported states:

| State | Site type | Notes |
|---|---|---|
| Virginia | Classic (requests) | |
| Massachusetts | Classic (requests) | |
| Colorado | Classic (requests) | |
| New Hampshire | Classic (requests) | |
| South Carolina | v2 (Playwright) | Requires Chromium |
| New Mexico | v2 (Playwright) | Requires Chromium |
| New York | v2 (Playwright) | Requires Chromium |

Classic states use simple HTTP requests and are fast. v2 states use a
headless Chromium browser (via Playwright) to render JavaScript — they are
slower but fully automated.

### Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | required | State name, snake\_case (e.g. `"virginia"`) |
| `year_from` | `1789` | Start year, inclusive |
| `year_to` | current year | End year, inclusive |
| `level` | `"all"` | What to return (see below) |
| `parallel` | `FALSE` | Parallel county scraping (classic states only) |

### The `level` parameter

| Value | Returns |
|---|---|
| `"all"` (default) | Named list with `$state` and `$county` data frames |
| `"state"` | Candidate-level results data frame |
| `"county"` | County vote breakdown data frame |
| `"joined"` | County rows merged with candidate metadata |

### Examples

```{r}
# Candidate-level results for Virginia, 2023
state_df <- scrape_elections(
  "election_stats",
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  level     = "state"
)

head(state_df)
```

```{r}
# County-level vote counts only
county_df <- scrape_elections(
  "election_stats",
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  level     = "county"
)
```

```{r}
# Both levels at once (default) — returns a named list
res <- scrape_elections(
  "election_stats",
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023
)

res$state   # candidate-level data frame
res$county  # county breakdown data frame
```

```{r}
# Multi-year range
ma_results <- scrape_elections(
  "election_stats",
  state     = "massachusetts",
  year_from = 2018,
  year_to   = 2022,
  level     = "state"
)
```

```{r}
# South Carolina (v2 / Playwright — slower but automatic)
sc_results <- scrape_elections(
  "election_stats",
  state     = "south_carolina",
  year_from = 2024,
  year_to   = 2024,
  level     = "state"
)
```

```{r}
# Parallel county scraping for classic states (6 workers)
res_parallel <- scrape_elections(
  "election_stats",
  state     = "virginia",
  year_from = 2023,
  year_to   = 2023,
  level     = "joined",
  parallel  = TRUE
)
```

### Columns returned — state level

The candidate-level (`"state"`) data frame typically includes:

- `state`, `year`, `election_id`, `candidate_id`
- `office`, `district`, `stage` (General, Primary, etc.)
- `candidate`, `party`
- `total_vote_count`, `vote_percentage`, `contest_outcome`
- `detail_url`

### Columns returned — county level

- `state`, `year`, `election_id`, `candidate_id`
- `county_or_city`, `candidate_name`, `votes`

---

## Source 2 — Ballotpedia (`source = "ballotpedia"`)

### What it covers

The Ballotpedia scraper collects school board election data from Ballotpedia's
annual school board election pages (2013 to present, all US states).

### Arguments

| Argument | Default | Description |
|---|---|---|
| `year` | `NULL` | Election year. Required for `mode = "results"` or `"joined"`. |
| `state` | `NULL` | State name (e.g. `"Alabama"`), or `NULL` for all states. |
| `mode` | `"districts"` | What to return (see below). |
| `start_year` | `2013` | Earliest year for multi-year district scrape. |
| `end_year` | `NULL` | Latest year (default: current calendar year). |

### The `mode` parameter

| Value | Returns | Speed |
|---|---|---|
| `"districts"` (default) | District-level metadata (one request per year-page) | Fast |
| `"results"` | Candidate names, votes, percentages per race | Slower — one extra request per district |
| `"joined"` | Districts + candidates in a single merged data frame | Slower |

### Examples

```{r}
# District metadata for all states, 2024 (fast — one page request)
districts_2024 <- scrape_elections(
  "ballotpedia",
  year = 2024
)

head(districts_2024)
```

```{r}
# Filter to one state
alabama_2024 <- scrape_elections(
  "ballotpedia",
  year  = 2024,
  state = "Alabama"
)
```

```{r}
# Full candidate results for one year + state (slower — follows each district)
results <- scrape_elections(
  "ballotpedia",
  year  = 2024,
  state = "New Jersey",
  mode  = "results"
)

# One row per candidate per race
head(results)
```

```{r}
# Districts + candidates joined into one table
joined <- scrape_elections(
  "ballotpedia",
  year  = 2024,
  state = "Texas",
  mode  = "joined"
)
```

```{r}
# Multi-year district metadata (no year argument, use start_year / end_year)
multi_year <- scrape_elections(
  "ballotpedia",
  state      = "Ohio",
  start_year = 2020,
  end_year   = 2024
)
```

### Columns returned — `mode = "districts"`

- `year`, `state`, `district`, `district_url`
- `primary`, `primary_runoff`, `general_election`, `general_runoff`
- `term_length`, `seats_up`, `total_board_seats`, `enrollment`

### Columns returned — `mode = "results"` / `"joined"`

All district columns plus:

- `race`, `election_type` (General / Primary / Primary Runoff / Other)
- `candidate`, `candidate_url`, `party`
- `is_winner`, `is_incumbent`
- `pct`, `votes`

---

## Source 3 — North Carolina Results (`source = "nc_results"`)

### What it covers

Returns local election results from the North Carolina State Board of
Elections for a specific election date.

### Arguments

| Argument | Default | Description |
|---|---|---|
| `date` | `NULL` | Election date string (e.g. `"2024-11-05"`), or `NULL` for all available. |

### Examples

```{r}
# All available NC election results
nc_all <- scrape_elections("nc_results")
```

```{r}
# Results for a specific election date
nc_nov2024 <- scrape_elections(
  "nc_results",
  date = "2024-11-05"
)

head(nc_nov2024)
```

---

## Working with the results

All sources return standard R `data.frame` objects (or a named list of them
when `level = "all"`). You can use any tidyverse tools on them directly:

```{r}
library(dplyr)

# ElectionStats: winners only, sorted by vote count
scrape_elections(
  "election_stats",
  state = "virginia", year_from = 2023, year_to = 2023, level = "state"
) |>
  filter(contest_outcome == "Winner") |>
  arrange(desc(total_vote_count)) |>
  select(office, district, candidate, party, total_vote_count)
```

```{r}
library(dplyr)

# Ballotpedia: competitive races (seats_up > 1)
scrape_elections("ballotpedia", year = 2024, state = "Texas") |>
  filter(as.integer(seats_up) > 1) |>
  select(district, general_election, seats_up, total_board_seats, enrollment)
```

---

## Choosing the right source

| Goal | Recommended source |
|---|---|
| Candidate + vote totals for VA, MA, CO, NH | `"election_stats"` |
| Candidate + vote totals for SC, NM, NY | `"election_stats"` (uses Playwright automatically) |
| School board district schedules, any state | `"ballotpedia"`, `mode = "districts"` |
| School board candidates + results | `"ballotpedia"`, `mode = "results"` |
| NC local election results | `"nc_results"` |

---

## Performance tips

- **Start small**: test with a single year and state before requesting large
  date ranges.
- **Playwright states are slower**: South Carolina, New Mexico, and New York
  launch a headless browser for each scrape. Expect several seconds per year.
- **Ballotpedia `mode = "results"`** makes one HTTP request per district;
  large states with many districts (e.g. Texas) can take several minutes.
- **`parallel = TRUE`** for ElectionStats speeds up county data collection on
  classic (requests-based) states by fetching detail pages concurrently.
- **Be polite**: the scrapers include built-in delays between requests. Do not
  disable them or reduce them aggressively — excessive requests may result in
  temporary IP blocks.
