---
title: "State elections (Ballotpedia)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{State elections (Ballotpedia)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

```{r setup}
library(DownBallotR)
library(dplyr)

downballot_use_python()
```

## What it covers

The Ballotpedia state elections scraper collects federal, state, and local
candidate listings from Ballotpedia's per-state election pages
(e.g. `https://ballotpedia.org/Maine_elections,_2024`). Use
`office = "state_elections"` to activate it.

Each page has three subsections — Federal Candidates, State Candidates,
Local Candidates — which you can filter with the `election_level` argument.

> **Coverage:** 2024 and 2025 page layouts are supported. Not all states
> have a Ballotpedia page for every year.

> **Vote counts** require `mode = "results"`, which follows each contest URL
> (one extra request per race). Use the default `mode = "listings"` for a
> fast candidate roster without vote data.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `office` | — | Must be `"state_elections"` |
| `state` | required | State name, title-case (e.g. `"Maine"`, `"Pennsylvania"`) |
| `year` | `NULL` | Single election year (e.g. `2024`). Required for `mode = "results"`. |
| `election_level` | `"all"` | Candidate tier to return — see below |
| `mode` | `"listings"` | What to return — see below |
| `start_year` | `2024` | Earliest year for multi-year listing scrape when `year` is `NULL` |
| `end_year` | `NULL` | Latest year (default: current calendar year) |

### The `election_level` parameter

| Value | Returns |
|---|---|
| `"all"` (default) | All candidate tiers |
| `"federal"` | U.S. House, U.S. Senate, Presidential Electors |
| `"state"` | Governor, state legislature, attorney general, etc. |
| `"local"` | Mayor, city council, county races, etc. |

### The `mode` parameter

| Value | Returns | Speed |
|---|---|---|
| `"listings"` (default) | One row per candidate from the state + year page; no vote counts | Fast — 1 request |
| `"results"` | All listing columns plus vote counts, percentages, and winner flags | Slower — 1 extra request per unique contest |

---

## Examples

### All candidates for a state — fast listing

```{r}
maine_2024 <- scrape_elections(
  state  = "Maine",
  office = "state_elections",
  year   = 2024
)

maine_2024 %>%
  select(level, office, jurisdiction, district, candidate, party, status) %>%
  arrange(level, office)
```

### Federal candidates only

```{r}
scrape_elections(
  state          = "Pennsylvania",
  office         = "state_elections",
  year           = 2024,
  election_level = "federal"
) %>%
  select(office, jurisdiction, district, candidate, party, status)
```

### Local candidates only

```{r}
scrape_elections(
  state          = "Maine",
  office         = "state_elections",
  year           = 2024,
  election_level = "local"
) %>%
  select(jurisdiction, office, district, candidate, party, status) %>%
  arrange(jurisdiction, office)
```

### Full results with vote counts

`mode = "results"` follows each contest link and returns votes and
winner/incumbent flags:

```{r}
maine_results <- scrape_elections(
  state  = "Maine",
  office = "state_elections",
  year   = 2024,
  mode   = "results"
)

# Winners only, sorted by vote share
maine_results %>%
  filter(is_winner) %>%
  arrange(desc(pct)) %>%
  select(level, office, jurisdiction, district, candidate, party, pct, votes)
```

### Local results with vote counts

```{r}
scrape_elections(
  state          = "Maine",
  office         = "state_elections",
  year           = 2024,
  election_level = "local",
  mode           = "results"
) %>%
  filter(is_winner) %>%
  select(jurisdiction, office, district, candidate, party, pct, votes) %>%
  arrange(jurisdiction, office)
```

### State-level incumbents running for re-election

```{r}
scrape_elections(
  state          = "Virginia",
  office         = "state_elections",
  year           = 2025,
  election_level = "state",
  mode           = "results"
) %>%
  filter(is_incumbent) %>%
  select(office, district, candidate, party, is_winner, pct, votes)
```

### Multi-year listing (2024–2025)

Use `start_year` / `end_year` when `year` is omitted:

```{r}
multi <- scrape_elections(
  state      = "Virginia",
  office     = "state_elections",
  start_year = 2024,
  end_year   = 2025
)

multi %>%
  group_by(year, level) %>%
  summarise(n_candidates = n(), .groups = "drop")
```

---

## Columns returned

### `mode = "listings"`

- `year`, `state`, `level` (federal / state / local)
- `contest_name` — raw office text as shown on Ballotpedia
- `contest_url` — link to the individual election page
- `jurisdiction` — parsed geographic area (e.g. `"Maine"`, `"Augusta"`)
- `office` — parsed role (e.g. `"U.S. Senate"`, `"City Council"`)
- `district` — parsed sub-race qualifier (e.g. `"District 1"`, `"At-large"`)
- `candidate`, `ballotpedia_url`, `party`, `status`
  (e.g. `"Won General"`, `"Lost Primary"`)

### `mode = "results"`

All listing columns plus:

- `election_type` (General / Primary / Primary Runoff / General (RCV) / Other)
- `is_winner`, `is_incumbent`
- `pct`, `votes`

\* Not all states have Ballotpedia state election pages for every year.
