---
title: "Utah elections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Utah elections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```
## What it covers

The Utah scraper returns election results from `electionresults.utah.gov`. It
covers all elections available on the site — generals, primaries, and special
elections — currently from 2023 to 2025.

Optionally, it can also return a vote-method breakdown (Advance in Person,
Election Day, Absentee by Mail, Provisional) for each contest.

> **Note:** The Utah elections site uses the same Angular/PrimeNG framework as
> the Georgia SOS site. All scraping uses a headless Chromium browser via
> Playwright. County scraping is parallelised across up to `max_workers`
> browsers (default 4). Utah has 29 counties.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | — | `"UT"`, `"utah"`, or `"Utah"` |
| `year_from` | `NULL` | Start year, inclusive; `NULL` = no lower bound |
| `year_to` | `NULL` | End year, inclusive; `NULL` = no upper bound |
| `level` | `"all"` | `"all"` — statewide + county; `"state"` — statewide only (much faster); `"county"` — county-level only |
| `max_workers` | `4L` | Parallel Chromium browsers for county scraping |
| `include_vote_methods` | `FALSE` | If `TRUE`, also return vote-method breakdown tables |

---

## Examples

### Statewide + county results for a single year

```{r, purl=FALSE}
res <- scrape_elections(state = "UT", year_from = 2024, year_to = 2024)

# Statewide candidate totals
res$state %>%
  select(election_name, office, candidate, party, votes, pct)

# County-level breakdown
res$county %>%
  select(county, office, candidate, party, votes, pct) %>%
  arrange(county, office, desc(votes))
```

### Statewide totals only (faster — skips county scraping)

```{r, purl=FALSE}
state_df <- scrape_elections(
  state     = "UT",
  year_from = 2024,
  year_to   = 2024,
  level     = "state"
)

state_df %>%
  select(election_name, office, candidate, party, votes, pct)
```

### With vote-method breakdown

```{r, purl=FALSE}
res <- scrape_elections(
  state                = "UT",
  year_from            = 2024,
  year_to              = 2024,
  include_vote_methods = TRUE
)

# Statewide vote-method totals
res$vote_method_state %>%
  select(office, candidate, party, votes_advance_in_person, votes_election_day,
         votes_absentee, votes_provisional, votes_total)

# County vote-method breakdown
res$vote_method_county %>%
  filter(county == "Salt Lake") %>%
  select(office, candidate, votes_advance_in_person, votes_election_day,
         votes_absentee, votes_provisional, votes_total)
```

### All available years

```{r, purl=FALSE}
res <- scrape_elections(
  state       = "UT",
  year_from   = 2023,
  year_to     = 2025,
  level       = "state"
)
```

### Filter to a specific county

```{r, purl=FALSE}
res <- scrape_elections(state = "UT", year_from = 2024, year_to = 2024)

res$county %>%
  filter(county == "Salt Lake") %>%
  select(election_name, office, candidate, party, votes, pct) %>%
  arrange(office, desc(votes))
```

---

## Output columns

### `$state` (when `level = "all"` or `"state"`)

| Column | Description |
|---|---|
| `election_name` | Human-readable election name |
| `election_year` | Calendar year of the election |
| `election_slug` | URL slug identifying the election |
| `election_date` | Date string from the page header |
| `result_status` | Reporting status (e.g. `"OFFICIAL RESULTS"`) |
| `office` | Office name |
| `localities_reporting` | `"X/Y"` localities reporting string |
| `candidate` | Candidate name (cleaned — `(I)` and party suffix stripped) |
| `party` | Party abbreviation |
| `is_incumbent` | `TRUE` if `(I)` appears in raw candidate name |
| `is_winner` | Always `NA` — winner marker absent from the site HTML |
| `votes` | Vote total |
| `pct` | Percentage of votes |
| `url` | URL of the election page |

### `$county` (when `level = "all"` or `"county"`)

Same columns as `$state`, plus:

| Column | Description |
|---|---|
| `county` | County name |

### `$vote_method_state` / `$vote_method_county` (when `include_vote_methods = TRUE`)

Same identifying columns as above, with votes split by method instead of a
single `votes` + `pct`:

| Column | Description |
|---|---|
| `votes_advance_in_person` | Advance in Person ballots |
| `votes_election_day` | Election Day ballots |
| `votes_absentee` | Absentee by Mail ballots |
| `votes_provisional` | Provisional ballots |
| `votes_total` | Sum of all methods |

---

## Performance notes

- **`level = "state"` is much faster.** County scraping opens up to
  `max_workers` parallel Chromium browsers across Utah's 29 counties.
- **`include_vote_methods` is slower.** Each contest panel requires an extra
  Playwright click-and-re-render step.
- **Virtual scroll.** The site uses Angular virtual scrolling; the scraper
  automatically scrolls to load all panels before parsing.
- **Limited history.** The Utah elections site currently lists elections from
  2023 onward. Earlier data is not available through this scraper.
