---
title: "Example Analyses with DownBallotR"
output: rmarkdown::html_vignette
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

`DownBallotR` returns standardized election results — candidates, parties,
votes, and winners — across federal, state, and local contests. This
vignette walks through several **example questions** users commonly want
to answer with the data, and shows the corresponding tidyverse code.

The examples use Indiana (`IN`) as a running example because Indiana
publishes partisan results for every office on the General Election
ballot and exposes both a statewide and a county-level frame. **Most
patterns transfer to other states**, but some details (exact office
strings, county column name) vary by source — each question below
calls out what's portable and what is not.

---

## A few things you might use this data for

- **Journalism** — Find the closest races of the cycle, identify which
  offices flipped party, or answer "how did each party do this year?"
- **Academic research** — Quantify down-ballot drop-off, study
  ticket-splitting, or assemble panel data on partisan races across years.
- **Local political strategy** — Understand the partisan baseline in a
  county, identify uncontested races, or target candidate recruitment
  in winnable districts.

---

## What varies across states (read this first)

A handful of details differ by source. They matter mostly when you start
filtering or cross-state binding:

- **Office strings are not normalized.** `office` is taken from each
  source's site, so what one source calls `"Governor"` another might
  call `"Governor of Indiana"` or `"GOVERNOR"`. Always inspect with
  `unique(df$office)` before filtering, and prefer `str_detect()` with
  a regex over `==`.
- **Subnational column names differ.** Indiana uses `county_name`,
  ElectionStats states (NH, MA, CO, ...) use `county_or_city`, Clarity
  states (GA, UT) use `county`, Connecticut uses `town`, Louisiana uses
  `parish`. See the
  [Data dictionary](https://gchickering21.github.io/DownBallotR/articles/data-dictionary.html)
  for the full schema per source.
- **Some columns only exist for some sources.** `election_date`,
  `election_id`, `num_seats`, `url`, `candidate_id` are source-specific.
  `bind_rows()` fills missing columns with `NA`, so cross-state binding
  still works — just don't reference those columns in cross-state code.
- **Party labels are normalized** to canonical full names
  (`"Democratic"`, `"Republican"`, `"Independent"`) across all sources,
  so cross-state party comparisons are safe.
- **`office_level`** is always one of `"Federal"`, `"State"`, or
  `"Local"`. Not every state populates all three.

---

## Setup

```{r}
library(DownBallotR)
library(tidyverse)

# Statewide + county results for Indiana, 2020–2024.
# When level = "all", scrape_elections() also assigns each sub-frame
# directly into your environment (e.g. in_state, in_county).
res <- scrape_elections(state = "IN", year_from = 2024, year_to = 2024)

# Either of these works — they refer to the same data frames:
in_state  <- res$state    # statewide candidate totals
in_county <- res$county   # county-level breakdown
```

A useful first step on any new scrape — see what office strings, office
levels, and parties are actually in the data:

```{r}
in_state %>% count(office_level, sort = TRUE)
in_state %>% distinct(office_level, office) %>% arrange(office_level, office)
in_state %>% count(party, sort = TRUE)
```

For reference, here's what Indiana actually returns for 2024:

```
# office_level counts (state-level frame)
office_level     n
Local         2316
State          306
Federal         51

# unique (office_level, office) — Federal & State
[Federal] Presidential Electors for US President & VP
[Federal] US Representative
[Federal] US Senator
[State]   Attorney General
[State]   Governor & Lt. Governor
[State]   Judge, Circuit Court / Probate Court / Superior Court
[State]   State Representative
[State]   State Senator

# parties present
Democratic, Republican, Libertarian, Independent,
Nonpartisan (1,100 rows — mostly judges and school boards), and a few rare labels.
```

Note two things up front:

- Indiana writes `"US Senator"` and `"US Representative"` (no periods).
  New Hampshire writes `"U.S. House"` and `"Governor"`. **Always inspect
  your data before filtering by exact office string.**
- Indiana labels many local races as `"Nonpartisan"` — judges, school
  boards. Filtering `party == "Democratic"` silently drops these from
  the denominator. For "partisan races only," filter
  `party %in% c("Democratic", "Republican", "Independent", "Libertarian")`
  or exclude `"Nonpartisan"`.

The remaining sections assume `in_state` and `in_county` are loaded.

---

## Question 1 — How many races did each party win?

Filter to winners and count by party:

```{r}
in_state %>%
  filter(winner) %>%
  count(party, sort = TRUE)
```

Often the more interesting view is by office level — federal, state, and
local performance can diverge:

```{r}
in_state %>%
  filter(winner) %>%
  count(office_level, party) %>%
  arrange(office_level, desc(n))
```

*Portability:* **fully portable.** `winner`, `party`, and `office_level`
are present on every state-level scrape.

---

## Question 2 — All races for a given office

The state-run General Election data covers every partisan office on the
ballot. To answer "how many of these races did each party win?" you
filter by office and count winners. Picking a single-word, universal
office like Governor keeps the regex simple:

```{r}
in_state %>%
  filter(str_detect(office, regex("governor", ignore_case = TRUE)))
```

`str_detect()` with a regex is more forgiving than `==`, since office
names vary slightly across sources and years (e.g. `"Governor"` vs
`"Governor of Indiana"` vs `"GOVERNOR"`).

Which party won each Governor's race in our window?

```{r}
in_state %>%
  filter(str_detect(office, regex("governor", ignore_case = TRUE)),
         winner) %>%
  select(election_year, office, candidate, party, votes, vote_pct)
```

The same pattern works for any office — swap `"governor"` for
`"senate"`, `"president"`, or `"sheriff"`. The user's original
"mayoral races by party" question is identical — `regex("mayor",
ignore_case = TRUE)` — though Indiana's General Election data does not
include municipal mayoral races (those run on odd years through
counties); Connecticut's data does.

*Portability:* **pattern is portable; the regex string is not.** Run
`unique(df$office)` first to confirm how your source spells the office.
A few real examples I observed:

- **U.S. House**: NH writes `"U.S. House"`, IN writes `"US Representative"`.
  Cross-state regex: `regex("u\\.?s\\.?\\s*(house|representative)", ignore_case = TRUE)`
- **U.S. Senate**: NH writes `"U.S. Senate"`, IN writes `"US Senator"`.
  Cross-state regex: `regex("senat", ignore_case = TRUE)` (matches both
  "Senate" and "Senator"; combine with `office_level == "Federal"` to
  exclude state senate).
- **Governor**: both NH and IN write something containing `"Governor"`
  as a substring (IN's full string is `"Governor & Lt. Governor"`), so
  `regex("governor", ignore_case = TRUE)` works for both.

---

## Question 3 — Total votes by party

Sometimes the question is not "who won" but "how much support did each
party receive overall?" Sum votes across all contests:

```{r}
in_state %>%
  group_by(party) %>%
  summarise(total_votes = sum(votes, na.rm = TRUE)) %>%
  arrange(desc(total_votes))
```

For a fairer comparison, restrict to a single office level — otherwise
a few high-turnout federal contests dominate the total:

```{r}
in_state %>%
  filter(office_level == "State") %>%
  group_by(election_year, party) %>%
  summarise(total_votes = sum(votes, na.rm = TRUE), .groups = "drop") %>%
  arrange(election_year, desc(total_votes))
```

*Portability:* **fully portable.**

---

## Question 4 — What were the closest races?

A common journalistic question. Compute each contest's winning margin in
percentage points by sorting candidates within a contest:

```{r}
in_state %>%
  group_by(election_year, office, district) %>%
  arrange(desc(vote_pct), .by_group = TRUE) %>%
  summarise(
    winner_party = first(party),
    winner_pct   = first(vote_pct),
    runner_pct   = nth(vote_pct, 2),
    margin_pp    = winner_pct - runner_pct,
    .groups      = "drop"
  ) %>%
  filter(!is.na(margin_pp)) %>%
  slice_min(margin_pp, n = 10)
```

`nth(vote_pct, 2)` returns `NA` for uncontested races, which the
`filter(!is.na(...))` step drops. `slice_min()` returns the 10 rows with
the smallest margin in ascending order.

*Portability:* **fully portable.** `(election_year, office, district)`
uniquely identifies a contest in every source schema.

---

## Question 5 — Down-ballot roll-off (a careful one)

Do voters who cast a top-of-ticket ballot also vote in down-ballot
races? This question is easy to ask but easy to get *wrong*.

**Tempting (but wrong) first attempt:** sum votes per office level.

```{r}
in_state %>%
  filter(election_year == 2024) %>%
  group_by(office_level) %>%
  summarise(total_votes = sum(votes, na.rm = TRUE)) %>%
  mutate(pct_of_federal = total_votes /
                          total_votes[office_level == "Federal"])
```

For Indiana 2024 this returns:

```
office_level total_votes pct_of_federal
Federal        8,634,681           1.00
State         12,420,340           1.44
Local         21,278,852           2.46
```

Local shows **2.46× as many votes as Federal**, which can't possibly
mean Local turnout was higher than presidential turnout. The reason:
each voter casts ballots in *several* local races (county treasurer +
county council + school board + …) but only a handful of federal ones
(president + their House district + maybe senator). So `sum(votes)`
counts `voters × (races per voter)`, and inflates by the number of
races at that level.

**The correct measure compares single races.** Each voter contributes
at most one vote to a given statewide race, so race-vs-race ratios
are real turnout comparisons:

```{r}
in_state %>%
  filter(election_year == 2024,
         office %in% c("Presidential Electors for US President & VP",
                       "Governor & Lt. Governor",
                       "US Senator",
                       "Attorney General")) %>%
  group_by(office) %>%
  summarise(total_votes = sum(votes, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(total_votes)) %>%
  mutate(pct_of_top = total_votes / max(total_votes))
```

Indiana 2024 actual output:

```
office                                       total_votes pct_of_top
Presidential Electors for US President & VP   2,936,677       1.000
Governor & Lt. Governor                       2,879,655       0.981
Attorney General                              2,838,098       0.966
US Senator                                    2,829,897       0.964
```

A few percentage points of roll-off down the statewide ballot is
typical — voters who skipped a race after voting for President.

**Local roll-off needs a different approach.** Each local race is
confined to a single county/district, so summing all 92 county
treasurer contests *does* equal ~one vote per voter (every voter has
one CT race available) — but only if every county actually holds the
race that year, which isn't always true. The safest local
roll-off measure: pick one county, then compare votes in its
top-of-ticket race to a specific local race within `in_county`.

```{r}
in_county %>%
  filter(election_year == 2024,
         county_name == "Marion",
         office %in% c("Presidential Electors for US President & VP",
                       "County Treasurer",
                       "School Board Member")) %>%
  group_by(office) %>%
  summarise(total_votes = sum(votes, na.rm = TRUE), .groups = "drop") %>%
  mutate(pct_of_president = total_votes /
            total_votes[office == "Presidential Electors for US President & VP"])
```

*Portability:* **the pattern is portable; the exact office strings
are state-specific.** Use `unique(df$office)` to find the top federal
race in your state. NH would use `"President"` and `"Governor"`, for
example, not Indiana's longer category labels.

---

## Question 6 — Year-over-year trends (update the initial data pull to ensure you have mutiple years of data)

Once you have a multi-year scrape, party performance over time is a
straightforward `count()`:

```{r}
in_state %>%
  filter(office_level == "Federal", winner) %>%
  count(election_year, party) %>%
  arrange(election_year, desc(n))
```

Or vote share for a single party over time:

```{r}
in_state %>%
  filter(office_level == "Federal") %>%
  group_by(election_year) %>%
  summarise(
    dem_share = sum(votes[party == "Democratic"], na.rm = TRUE) /
                sum(votes, na.rm = TRUE)
  )
```

*Portability:* **fully portable.**

---

## Question 7 — County-level variation within a single race

Where did a candidate over- or under-perform their statewide result?
Using the county-level frame, look at Democratic vote share for the
2024 U.S. Senate race across Indiana's 92 counties.

Note the regex below uses `"senat"`, not `"senate"` — Indiana writes
the office as `"US Senator"`, so `regex("senate", ...)` would return
zero rows. `"senat"` matches both `"Senate"` (NH) and `"Senator"` (IN),
and the `office_level == "Federal"` filter excludes State Senate.

```{r}
in_county %>%
  filter(election_year == 2024,
         office_level == "Federal",
         str_detect(office, regex("senat", ignore_case = TRUE)),
         party == "Democratic") %>%
  arrange(desc(vote_pct)) %>%
  select(county_name, candidate, votes, vote_pct)
```

To find the counties where the statewide nominee under-performed most,
pull the statewide number and compute a swing for each county:

```{r}
statewide_dem_senate <- in_state %>%
  filter(election_year == 2024,
         office_level == "Federal",
         str_detect(office, regex("senat", ignore_case = TRUE)),
         party == "Democratic") %>%
  pull(vote_pct)

in_county %>%
  filter(election_year == 2024,
         office_level == "Federal",
         str_detect(office, regex("senat", ignore_case = TRUE)),
         party == "Democratic") %>%
  mutate(swing_vs_state = vote_pct - statewide_dem_senate) %>%
  select(county_name, vote_pct, swing_vs_state) %>%
  slice_min(swing_vs_state, n = 10)
```

*Portability:* **pattern is portable; the column name is not.** The
subnational unit column changes by source:

| Source | Subnational column |
|---|---|
| Indiana | `county_name` |
| ElectionStats (NH, MA, CO, …) | `county_or_city` |
| Clarity (GA, UT) | `county` |
| Connecticut | `town` |
| Louisiana | `parish` |
| North Carolina | `county` (county frame) or `precinct` |

---

## Drilling into local elections

State-run General Election scrapers include a long tail of local offices
alongside the federal and statewide races. Indiana 2024 alone returns
**2,316 local-office rows** across 14 office types (school boards,
county treasurers, county commissioners, town councils, township boards,
etc.). Several questions are only possible at this level.

**First, a data-shape caveat.** For local offices in Indiana's
state-level frame, the `district` column means different things
depending on the office:

- *Countywide offices* (County Treasurer, County Recorder, County
  Auditor, etc.) have no sub-district, so the parser puts the **county
  name** into `district` (e.g. `"Bartholomew County"`). One row of
  `(office, district)` = one contest.
- *Districted offices* (County Council, County Commissioner) have an
  internal district. The parser puts the **district label** into
  `district` (e.g. `"District 1"`), and the county name is **lost from
  the state frame**. To analyze these by county, use the county-level
  frame (`in_county`), which has `county_name` as a separate column.

The examples below stick to offices where `district` reliably encodes
the contest.

### What local offices are on the ballot, and which are partisan?

```{r}
in_state %>%
  filter(office_level == "Local") %>%
  group_by(office) %>%
  summarise(
    n_candidates = n_distinct(candidate),
    is_partisan  = any(party %in% c("Democratic", "Republican")),
    parties      = paste(sort(unique(party)), collapse = ", "),
    .groups      = "drop"
  ) %>%
  arrange(desc(n_candidates))
```

Real output for Indiana 2024 (abbreviated):

```
office                       n_candidates  is_partisan
School Board Member                  1086        FALSE
County Council Member                 408         TRUE
County Commissioner                   262         TRUE
Town Council Member                   125         TRUE
County Treasurer                       88         TRUE
County Coroner                         86         TRUE
County Surveyor                        71         TRUE
Township Board Member                  57         TRUE
County Recorder                        47         TRUE
Clerk of the Circuit Court             34         TRUE
County Auditor                         28         TRUE
Town Clerk-Treasurer                   21         TRUE
```

Useful as a first orientation: nine of the twelve most common local
office types are partisan, with school boards (the single largest local
office by candidate count) the major nonpartisan exception.

### How competitive are countywide partisan races?

County Treasurer is a textbook countywide partisan office — every
county that holds the election has exactly one contest, so
`(office, district)` cleanly identifies each race. The natural
question: how often is the race contested, and how often is it
one-party?

```{r}
in_state %>%
  filter(office == "County Treasurer",
         party %in% c("Democratic", "Republican")) %>%
  group_by(district, election_year) %>%      # district = county name
  summarise(
    n_candidates = n_distinct(candidate),
    parties      = paste(sort(unique(party)), collapse = "/"),
    .groups      = "drop"
  ) %>%
  count(n_candidates, parties)
```

The output is a small table — for Indiana 2024 you'll typically see
many `n_candidates == 1` rows (single party fielding the only
candidate) alongside the `Democratic/Republican` two-candidate races.
The same pattern works for County Recorder, County Auditor, and other
countywide partisan offices.

For an explicitly partisan-baseline view, group by the winner's party:

```{r}
in_state %>%
  filter(office == "County Treasurer",
         party %in% c("Democratic", "Republican"),
         winner) %>%
  count(party)
```

This is the local analogue of "how many county treasurer races did each
party win?" — the user's original mayoral question, restated for an
office that actually appears in this dataset.

### Are school board races competitive?

School Board Member is by far the largest local office in Indiana's
data (~1,000 candidates per cycle across ~300 contests). The
distribution of candidates per contest measures field size:

```{r}
in_state %>%
  filter(office == "School Board Member") %>%
  group_by(district, election_year) %>%
  summarise(n_candidates = n_distinct(candidate), .groups = "drop") %>%
  count(n_candidates) %>%
  arrange(n_candidates)
```

Most school board contests draw 1–4 candidates; a long right tail of
large at-large boards (10+, sometimes 50+ candidates) reflects
multi-seat elections rather than head-to-head races.

*Portability:* **patterns are portable; office strings are not.** The
set of local offices a source reports varies a lot — Indiana exposes
14 distinct local office types; NH exposes 6 (Attorney, County
Commissioner, Register of Deeds, Register of Probate, Sheriff,
Treasurer); Connecticut surfaces actual municipal mayors that Indiana
omits. Always inspect `unique(df$office)` for your state first, and
prefer the county/town/parish frame for cross-jurisdictional analysis
of districted offices.

---

## Across multiple states (small N)

`scrape_elections()` returns one state at a time, but the output schemas
are consistent enough that you can bind frames together and compare. The
`state` column is always populated, so cross-state queries work cleanly
even when the column sets differ slightly. `bind_rows()` fills `NA`
where columns aren't shared.

### Pulling several states

```{r}
ind <- scrape_elections(state = "IN", year_from = 2024, year_to = 2024,
                        level = "state")
nh  <- scrape_elections(state = "NH", year_from = 2024, year_to = 2024,
                        level = "state")

multi <- bind_rows(ind, nh)
```

### Question 8 — Governor winners across states

Which party won the Governor's race in each state? (Both IN and NH held
gubernatorial elections in 2024.)

```{r}
multi %>%
  filter(str_detect(office, regex("governor", ignore_case = TRUE)),
         winner) %>%
  select(state, candidate, party, votes, vote_pct)
```

### Question 9 — Democratic vote share across states

A single statewide number per state — useful for ranking how each state
voted in 2024 at the federal level:

```{r}
multi %>%
  filter(office_level == "Federal") %>%
  group_by(state) %>%
  summarise(
    dem_share = sum(votes[party == "Democratic"], na.rm = TRUE) /
                sum(votes, na.rm = TRUE),
    rep_share = sum(votes[party == "Republican"], na.rm = TRUE) /
                sum(votes, na.rm = TRUE)
  ) %>%
  arrange(desc(dem_share))
```

### Question 10 — Down-ballot roll-off, state by state

Applying the **correct** Question 5 logic — single top race at each
level — to multiple states puts them on a comparable scale. For each
state, find the highest-vote race at the Federal and State levels,
then compute the ratio:

```{r}
multi %>%
  filter(office_level %in% c("Federal", "State")) %>%
  group_by(state, office_level, office) %>%
  summarise(office_votes = sum(votes, na.rm = TRUE), .groups = "drop") %>%
  group_by(state, office_level) %>%
  slice_max(office_votes, n = 1, with_ties = FALSE) %>%   # top race per level
  group_by(state) %>%
  mutate(pct_of_top_federal = office_votes /
                              office_votes[office_level == "Federal"]) %>%
  select(state, office_level, office, office_votes, pct_of_top_federal) %>%
  arrange(state, desc(office_votes))
```

The Federal row is 1.00 by construction; the State row's
`pct_of_top_federal` reveals how much turnout dropped between the top
federal race and the top state race in each state. (For local
roll-off, see the note in Question 5 — use a single county's data
from the county-level frame.)

### A note on `vote_pct`

`vote_pct` is normalized within each contest, so it's *not* directly
comparable across contests or states. For cross-state comparisons,
always work from raw `votes` and re-aggregate (as in Questions 9 and 10
above).

### A note on party labels in cross-state queries

Most sources use canonical `"Democratic"` / `"Republican"` labels, but
a few quirks show up when binding many states:

- **New Hampshire** has fusion-party labels — candidates appear with
  `"Republican/Democratic"` or `"Democratic/Republican"` when nominated
  by both major parties. A filter `party == "Democratic"` silently
  drops them. Use `str_detect(party, "Democratic")` if you want to
  count fusion candidates as Democrats.
- **Indiana** labels many local races (judges, school boards) as
  `"Nonpartisan"`. These have no Dem/Rep candidate at all. Include
  `filter(party != "Nonpartisan")` if your question is about partisan
  competition only.

---

## Comparing across many states (large N)

Once you have data for a dozen or more states, the questions shift from
"who won" to "how do states differ on this metric?" The pattern is
always the same: scrape each state once, bind them, then compute one
number per state. Below are questions that only become interesting at
scale.

### Building a multi-state panel

Scraping many states in one R session can take a while. A common
pattern is to scrape each state once, save the result, and re-load
later:

```{r}
states <- c("IN", "NH", "VA", "MA", "CO", "VT", "ID", "NY", "NM", "SC")

# Pull each state's 2024 state-level frame, bind, and save
panel_2024 <- map_dfr(states, function(s) {
  scrape_elections(state = s, year_from = 2024, year_to = 2024,
                   level = "state")
})
saveRDS(panel_2024, "panel_2024.rds")

# Next session — skip the rescrape
panel_2024 <- readRDS("panel_2024.rds")
```

`map_dfr()` (from purrr) calls `scrape_elections()` once per state and
row-binds the results. If one state errors, wrap the call in
`possibly()` or `safely()` so a single failure doesn't lose the rest.

### Question 11 — Closest statewide race in each state

Which state had the nail-biter of the cycle? Combine Q4's margin
calculation with a `group_by(state)`:

```{r}
panel_2024 %>%
  filter(office_level == "Federal") %>%
  group_by(state, office, district) %>%
  arrange(desc(vote_pct), .by_group = TRUE) %>%
  summarise(
    winner_party = first(party),
    margin_pp    = first(vote_pct) - nth(vote_pct, 2),
    .groups      = "drop"
  ) %>%
  filter(!is.na(margin_pp)) %>%
  group_by(state) %>%
  slice_min(margin_pp, n = 1) %>%
  arrange(margin_pp)
```

This returns the single tightest federal race per state, ranked
nationally — a useful one-glance summary of where the action was.

### Question 12 — Two-party vote share, ranked across states

How does Democratic vote share for federal races compare across the
panel?

```{r}
panel_2024 %>%
  filter(office_level == "Federal") %>%
  group_by(state) %>%
  summarise(
    n_contests = n_distinct(office, district),
    dem_share  = sum(votes[party == "Democratic"], na.rm = TRUE) /
                 sum(votes, na.rm = TRUE),
    rep_share  = sum(votes[party == "Republican"], na.rm = TRUE) /
                 sum(votes, na.rm = TRUE)
  ) %>%
  mutate(two_party_dem = dem_share / (dem_share + rep_share)) %>%
  arrange(desc(two_party_dem))
```

`two_party_dem` puts states on a comparable scale even when third-party
share varies.

### Question 13 — Uncontested races by state

Which states have the most one-candidate races? A useful structural
indicator of party competition:

```{r}
panel_2024 %>%
  group_by(state, office_level, office, district, election_year) %>%
  summarise(n_candidates = n_distinct(candidate), .groups = "drop") %>%
  filter(n_candidates == 1) %>%
  count(state, office_level, sort = TRUE)
```

Or as a rate — share of contests that are uncontested:

```{r}
panel_2024 %>%
  group_by(state, office_level, office, district, election_year) %>%
  summarise(n_candidates = n_distinct(candidate), .groups = "drop") %>%
  group_by(state, office_level) %>%
  summarise(
    n_contests   = n(),
    n_uncontested = sum(n_candidates == 1),
    pct_uncontested = n_uncontested / n_contests,
    .groups = "drop"
  ) %>%
  arrange(desc(pct_uncontested))
```

### Question 14 — Down-ballot roll-off, ranked nationally

Q10 with more states — the answers become a national distribution
rather than a two-state comparison. Using the same "single top race
per level" approach to avoid the per-voter-multiplier bug in
Question 5:

```{r}
panel_2024 %>%
  filter(office_level %in% c("Federal", "State")) %>%
  group_by(state, office_level, office) %>%
  summarise(office_votes = sum(votes, na.rm = TRUE), .groups = "drop") %>%
  group_by(state, office_level) %>%
  slice_max(office_votes, n = 1, with_ties = FALSE) %>%
  group_by(state) %>%
  summarise(
    top_federal_votes = office_votes[office_level == "Federal"],
    top_state_votes   = office_votes[office_level == "State"],
    state_to_federal  = top_state_votes / top_federal_votes,
    .groups = "drop"
  ) %>%
  arrange(state_to_federal)
```

States at the top of the result lose the most voters between the
top of the ticket and the top statewide office (e.g., Governor or
Attorney General) — a useful indicator of how much voter attention
extends past President. For local-level roll-off across many states,
loop over a single county per state using `in_county` (the local
comparison only works when every voter shares the race, which doesn't
hold across multiple counties).

### Question 15 — Party dominance index

A composite of "share of statewide federal contests won by the dominant
party," useful for ranking states by partisan asymmetry:

```{r}
panel_2024 %>%
  filter(office_level == "Federal", winner) %>%
  count(state, party) %>%
  group_by(state) %>%
  summarise(
    total_wins   = sum(n),
    top_party    = party[which.max(n)],
    top_share    = max(n) / sum(n),
    .groups      = "drop"
  ) %>%
  arrange(desc(top_share))
```

---

## Quick top-line: `summarize_results()`

Before writing targeted queries, `summarize_results()` gives a one-call
overview of what's in any scraped data frame:

```{r}
summarize_results(in_state)
```

Output (abbreviated):

```
Election Results Summary
========================
State            : Indiana
Years covered    : 2020–2024  (3 years)
Elections        : ...
Unique candidates: ...

Elections by office level:
  Federal  : N elections, M offices
             - ...
  State    : ...
  Local    : ...
```

It is a useful starting point for orienting yourself before drilling
into a specific office, party, or year.
