---
title: "Georgia elections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Georgia elections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```
## What it covers

The Georgia scraper returns election results from the Georgia Secretary of State
at `results.sos.ga.gov`. It covers all elections available on the site —
primaries, runoffs, generals, and special elections — from 2000 to the present.

Results are available at three levels: statewide candidate totals, county
breakdowns, and precinct-level results. Precinct data is obtained by navigating
each county page and following "View results by precinct" links for each contest.

Optionally, it can also return a vote-method breakdown (Advance in Person,
Election Day, Absentee by Mail, Provisional) for each contest.

> **Note:** The Georgia SOS site is a JavaScript-rendered Angular SPA. All
> scraping uses a headless Chromium browser via Playwright. County and precinct
> scraping is parallelised across up to `max_workers` browsers (default 4).
> Georgia has 159 counties; a full state + county scrape for one year typically
> takes 20–60 minutes. Adding precinct data significantly increases runtime.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | — | `"GA"`, `"georgia"`, or `"Georgia"` |
| `year_from` | `NULL` | Start year, inclusive; `NULL` = no lower bound |
| `year_to` | `NULL` | End year, inclusive; `NULL` = no upper bound |
| `level` | `"all"` | `"all"` — statewide + county + precinct; `"state"` — statewide only (fastest); `"county"` — county-level only; `"precinct"` — precinct-level only |
| `max_workers` | `4L` | Parallel Chromium browsers for county and precinct scraping |
| `include_vote_methods` | `FALSE` | If `TRUE`, also return vote-method breakdown tables |

---

## Examples

### Statewide + county results for a single year

```{r, purl=FALSE}
res <- scrape_elections(state = "GA", year_from = 2024, year_to = 2024)

# Statewide candidate totals
res$state %>%
  filter(office == "President of the United States") %>%
  select(election_name, candidate, party, votes, vote_pct)

# County-level breakdown
res$county %>%
  filter(office == "President of the United States") %>%
  select(county, candidate, party, votes, vote_pct) %>%
  arrange(county, desc(votes))
```

### Statewide totals only (faster — skips county scraping)

```{r, purl=FALSE}
state_df <- scrape_elections(
  state     = "GA",
  year_from = 2024,
  year_to   = 2024,
  level     = "state"
)

state_df %>%
  select(election_name, office, candidate, party, votes, vote_pct)
```

### With vote-method breakdown

```{r, purl=FALSE}
res <- scrape_elections(
  state                = "GA",
  year_from            = 2024,
  year_to              = 2024,
  include_vote_methods = TRUE
)

# Statewide vote-method totals
res$vote_method_state %>%
  filter(office == "United States Senator") %>%
  select(candidate, party, votes_advance_voting, votes_election_day,
         votes_absentee, votes_provisional, votes_total)

# County vote-method breakdown
res$vote_method_county %>%
  filter(office == "United States Senator", county == "Fulton") %>%
  select(candidate, votes_advance_voting, votes_election_day,
         votes_absentee, votes_provisional, votes_total)
```

### Faster county scraping with more parallel browsers

Georgia has 159 counties. By default, county pages are scraped with 4 parallel
Chromium browsers. Increasing `max_workers` can cut runtime significantly on
machines with enough CPU/RAM.

```{r, purl=FALSE}
res <- scrape_elections(
  state       = "GA",
  year_from   = 2024,
  year_to     = 2024,
  level       = "all",
  max_workers = 8L
)
```

### Multi-year scrape

```{r, purl=FALSE}
res <- scrape_elections(
  state       = "GA",
  year_from   = 2020,
  year_to     = 2024,
  level       = "all",
  max_workers = 4L
)
```

### Precinct-level results

```{r, purl=FALSE}
# level = "all" includes precinct data automatically
res <- scrape_elections(state = "GA", year_from = 2024, year_to = 2024)
res$precinct %>%
  filter(office == "President of the United States", county == "Fulton") %>%
  select(precinct, candidate, party, votes, vote_pct) %>%
  arrange(precinct, desc(votes))

# Precinct only (skips statewide + county scraping)
precinct_df <- scrape_elections(
  state     = "GA",
  year_from = 2024,
  year_to   = 2024,
  level     = "precinct"
)
```

### Filter to a specific county

```{r, purl=FALSE}
res <- scrape_elections(state = "GA", year_from = 2022, year_to = 2022)

res$county %>%
  filter(county == "Fulton") %>%
  select(election_name, office, candidate, party, votes, vote_pct) %>%
  arrange(office, desc(votes))
```

---

## Output columns

### `$state` (when `level = "all"` or `"state"`)

| Column | Description |
|---|---|
| `state` | State abbreviation (`"GA"`) |
| `election_name` | Human-readable election name (e.g. `"November General Election"`) |
| `election_type` | Election type (e.g. `"General"`, `"Primary"`) |
| `election_year` | Calendar year of the election |
| `election_date` | Date string from the page header |
| `office_level` | `"Federal"`, `"State"`, or `"Local"` |
| `office` | Office name |
| `district` | District identifier; `NA` when not applicable |
| `candidate` | Candidate name |
| `party` | Party abbreviation |
| `votes` | Vote total |
| `vote_pct` | Percentage of votes within the contest |
| `winner` | `TRUE` if the candidate won the statewide contest |
| `url` | URL of the election results page scraped |

### `$county` (when `level = "all"` or `"county"`)

| Column | Description |
|---|---|
| `state` | State abbreviation (`"GA"`) |
| `election_name` | Human-readable election name |
| `election_type` | Election type (e.g. `"General"`, `"Primary"`) |
| `election_year` | Calendar year of the election |
| `election_date` | Date string from the page header |
| `office_level` | `"Federal"`, `"State"`, or `"Local"` |
| `office` | Office name |
| `district` | District identifier; `NA` when not applicable |
| `county` | County name |
| `candidate` | Candidate name |
| `party` | Party abbreviation |
| `votes` | County-level vote total |
| `vote_pct` | Percentage of votes within the contest for this county |
| `county_winner` | `TRUE` if the candidate won within this county |
| `url` | URL of the election results page scraped |

### `$precinct` (when `level = "all"` or `"precinct"`)

| Column | Description |
|---|---|
| `state` | State abbreviation (`"GA"`) |
| `election_name` | Human-readable election name |
| `election_type` | Election type (e.g. `"General"`, `"Primary"`) |
| `election_year` | Calendar year of the election |
| `election_date` | Date string from the page header |
| `office_level` | `"Federal"`, `"State"`, or `"Local"` |
| `office` | Office name |
| `district` | District identifier; `NA` when not applicable |
| `county` | County name |
| `precinct` | Precinct name |
| `candidate` | Candidate name |
| `party` | Party abbreviation |
| `votes` | Precinct-level vote total |
| `vote_pct` | Percentage of votes within the contest for this precinct |
| `precinct_winner` | `TRUE` if the candidate won within this precinct |
| `url` | URL of the precinct results page scraped |

### `$vote_method_state` / `$vote_method_county` (when `include_vote_methods = TRUE`)

Same identifying columns as above, with votes split by method instead of a
single `votes` + `pct`:

| Column | Description |
|---|---|
| `votes_advance_voting` | Advance in Person ballots |
| `votes_election_day` | Election Day ballots |
| `votes_absentee` | Absentee by Mail ballots |
| `votes_provisional` | Provisional ballots |
| `votes_total` | Sum of all methods |

---

## Performance notes

- **`level = "state"` is much faster.** County scraping opens up to
  `max_workers` parallel Chromium browsers across Georgia's 159 counties.
  A full state + county scrape for one year takes 20–60 minutes. Use
  `level = "state"` when county breakdowns are not needed.
- **Precinct scraping is significantly slower.** It navigates each county page,
  fetches one ballot-item page per contest (in parallel), and then scrapes each
  individual precinct page. Expect a full state + county + precinct run to take
  several hours for a general election year. Use `level = "county"` or
  `level = "state"` when precinct data is not needed.
- **`include_vote_methods` is significantly slower.** Each contest panel requires
  an extra Playwright click-and-re-render step. Budget an extra 5–15 minutes per
  election page.
- **Virtual scroll.** The GA SOS site uses Angular virtual scrolling, rendering
  only panels visible in the viewport. The scraper automatically scrolls to load
  all panels before parsing — without this, pages with 100+ contests would
  silently return incomplete data.
- **Multiple elections per year.** The GA SOS site lists all elections for a year
  (primaries, runoffs, generals, specials). All are scraped and combined in the
  output.
