---
title: "Louisiana elections"
output: rmarkdown::html_vignette
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```
## What it covers

The Louisiana scraper returns election results from the Louisiana Secretary of
State Graphical results site at `voterportal.sos.la.gov`. It covers all
elections listed in the site's dropdown — primaries, runoffs, generals, and
special elections — from 1982 to the present.

Results are scraped from non-Parish tabs (Congressional, Statewide, Legislative,
Multiparish) for statewide totals, and from the Parish tab for parish-level
breakdowns. Voter turnout percentages are captured at the page level when
published by the site.

> **Note:** The Louisiana SOS site is a JavaScript-rendered AngularJS
> application. All scraping uses a headless Chromium browser via Playwright.
> Parish scraping is parallelised across up to `max_workers` browsers
> (default 2). Louisiana has 64 parishes.

---

## Arguments

| Argument | Default | Description |
|---|---|---|
| `state` | — | `"LA"`, `"louisiana"`, or `"Louisiana"` |
| `year_from` | `NULL` | Start year, inclusive; `NULL` = no lower bound (data goes back to 1982) |
| `year_to` | `NULL` | End year, inclusive; `NULL` = current calendar year |
| `level` | `"all"` | `"all"` — statewide + parish; `"state"` — statewide only (much faster); `"parish"` — parish-level only |
| `max_workers` | `4L` | Parallel Chromium browsers for parish scraping (capped at 2 internally for LA) |

---

## Examples

### Statewide + parish results for a single year

```{r, purl=FALSE}
res <- scrape_elections(state = "LA", year_from = 2024, year_to = 2024)

# Statewide candidate totals
res$state %>%
  filter(office_level == "Federal") %>%
  select(election_name, office, candidate, party, votes, vote_pct, winner)

# Parish-level breakdown
res$parish %>%
  filter(office == "President of the United States") %>%
  select(parish, candidate, party, votes, vote_pct) %>%
  arrange(parish, desc(votes))
```

### Statewide totals only (faster — skips parish scraping)

```{r, purl=FALSE}
state_df <- scrape_elections(
  state     = "LA",
  year_from = 2024,
  year_to   = 2024,
  level     = "state"
)

state_df %>%
  select(election_name, office_level, office, candidate, party, votes, vote_pct, winner)
```

### Multi-year scrape

```{r, purl=FALSE}
res <- scrape_elections(
  state     = "LA",
  year_from = 2020,
  year_to   = 2024,
  level     = "state"
)

res %>%
  filter(office == "President of the United States") %>%
  select(election_name, election_year, candidate, party, votes, vote_pct, winner)
```

### Filter to a specific parish

```{r, purl=FALSE}
res <- scrape_elections(state = "LA", year_from = 2023, year_to = 2023)

res$parish %>%
  filter(parish == "Orleans Parish") %>%
  select(election_name, office, candidate, party, votes, vote_pct) %>%
  arrange(office, desc(votes))
```

### Filter by office level

```{r, purl=FALSE}
res <- scrape_elections(state = "LA", year_from = 2024, year_to = 2024)

# Federal races only
res$state %>%
  filter(office_level == "Federal") %>%
  select(election_name, office, candidate, party, votes, vote_pct, winner)

# State-level races
res$state %>%
  filter(office_level == "State") %>%
  select(election_name, office, candidate, party, votes, vote_pct, winner)
```

---

## Output columns

### `$state` (when `level = "all"` or `"state"`)

| Column | Description |
|---|---|
| `state` | State abbreviation (`"LA"`) |
| `election_name` | Human-readable election name (e.g. `"November 5, 2024 General Election"`) |
| `election_year` | Calendar year of the election |
| `election_date` | Date of the election |
| `office_level` | `"Federal"`, `"State"`, or `"Local"` |
| `office` | Office title |
| `district` | Court, district, or seat qualifier; `NA` when not present |
| `candidate` | Candidate name |
| `party` | Full party name (e.g. `"Republican"`, `"Democratic"`) |
| `votes` | Vote total |
| `vote_pct` | Percentage of votes within the contest |
| `winner` | `TRUE` if the candidate won |
| `voter_turnout_pct` | Statewide voter turnout percentage; `NA` when not published |

### `$parish` (when `level = "all"` or `"parish"`)

| Column | Description |
|---|---|
| `state` | State abbreviation (`"LA"`) |
| `election_name` | Human-readable election name |
| `election_year` | Calendar year of the election |
| `election_date` | Date of the election |
| `office_level` | `"Federal"`, `"State"`, or `"Local"` |
| `office` | Office title |
| `district` | Court, district, or seat qualifier; `NA` when not present |
| `parish` | Parish name (e.g. `"Orleans Parish"`) |
| `candidate` | Candidate name |
| `party` | Full party name |
| `votes` | Parish-level vote total |
| `vote_pct` | Percentage of votes within the contest for this parish |
| `parish_winner` | `TRUE` if the candidate won within this parish |
| `parish_voter_turnout_pct` | Parish-level voter turnout percentage; `NA` when not published |

---

## Performance notes

- **`level = "state"` is much faster.** Parish scraping opens up to
  `max_workers` parallel Chromium browsers across Louisiana's 64 parishes.
  A full state + parish scrape for one year typically takes 15–40 minutes.
  Use `level = "state"` when parish breakdowns are not needed.
- **Multiple elections per year.** The Louisiana SOS site lists all elections
  for a given year (primaries, runoffs, generals, specials). All are scraped
  and combined in the output.
- **Cross-tab deduplication.** Some races (e.g. Multiparish offices) appear
  in multiple tabs. The scraper automatically deduplicates these, keeping the
  highest-priority tab's data.
- **Long history.** The dropdown goes back to 1982; however, older elections
  may have sparser data or slightly different page layouts.
- **Turnout data.** The `voter_turnout_pct` and `parish_voter_turnout_pct`
  columns are `NA` for elections where the site does not publish turnout.
