---
title: "Python setup for DownBallotR"
output: rmarkdown::html_vignette
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## 1. Install the package

You can install the development version from GitHub:

```{r, purl=FALSE}
install.packages("pak")
pak::pak("gchickering21/Downballot")
```


Or using `remotes`:

```{r, purl=FALSE}
install.packages("remotes")
remotes::install_github("gchickering21/Downballot")
```

Then load the package:

```{r, purl=FALSE}
library(DownBallotR)
```


## 2. First-time setup (required once)

After installing the R package, you must set up the Python environment used by
`DownBallotR`.

### Step 1: Install Python via reticulate

`DownBallotR` manages its own isolated Python environment through `reticulate`.
**You do not need to install Python yourself** — the following command downloads
and installs a standalone Python that reticulate manages independently of
anything already on your machine:

```{r, purl=FALSE}
reticulate::install_python()
```

This works on Windows, macOS, and Linux without any prior Python installation.
You can pin a specific version if needed (e.g.
`reticulate::install_python(version = "3.12")`); any version 3.10 or later works.

### Step 2: Install Python dependencies

Once Python is available, run the following **once**:

```{r, purl=FALSE}
downballot_install_python(python = reticulate::virtualenv_starter())
```

`reticulate::virtualenv_starter()` automatically selects the best available
Python for creating a virtual environment — matching whatever version you just
installed. If you installed a specific version in Step 1, pass the same version
here (e.g. `reticulate::virtualenv_starter("3.12")`).

This command:

- Creates a dedicated Python virtual environment for `DownBallotR`
- Installs required Python packages
- Installs Playwright and its Chromium browser
  (this may download ~100–200MB the first time)

This step may take a few minutes.

> **Note:** If you already have Python 3.10 or later installed and working on
> your machine, you can try `downballot_install_python()` without arguments.
> If that fails (e.g. with a virtualenv creation error), use the
> `reticulate::install_python()` approach above.

---

### Step 2: Verify the setup

After installation completes, check the status:

```{r, purl=FALSE}
downballot_python_status()
```

You should see output indicating:

- The virtual environment exists
- Required Python packages are installed
- Playwright Chromium is available

At this point, Python is installed but **not yet active** in your R session.

---

## 3. Normal usage (each R session)

Each R session must explicitly activate the Python environment **once per
session**.

### Activate Python for the session

At the start of any R session where you plan to use `DownBallotR`, run:

```{r, purl=FALSE}
downballot_use_python()
```

This:

- Pins `reticulate` to the correct Python virtual environment
- Safely initializes Python for the current R session

You only need to run this once per session.

---

### Optional: check status

At any time, you can check the current state of the Python environment:

```{r, purl=FALSE}
downballot_python_status()
```

This reports:

- Whether Python is initialized
- Which Python interpreter is active
- Whether required packages are available

---

## 4. Using DownBallotR in future or concurrent sessions

### New R session?

If you restart R, open a new RStudio window, or start a separate R process,
simply run:

```{r, purl=FALSE}
downballot_use_python()
```


You **do not** need to reinstall Python.

---

### Multiple R sessions at the same time

If you have multiple R sessions open concurrently:

- Each session must call `downballot_use_python()` once
- All sessions can safely share the same virtual environment

---

## 5. Platform notes

### Windows

The setup works on Windows without modification. When installing Python via
`reticulate::install_python()`, ensure you run R with normal user permissions
(administrator is not required). Playwright downloads a self-contained Chromium
bundle on Windows, so no additional system software is needed.

### macOS

The recommended `reticulate::install_python()` approach works on macOS. If you
prefer to use a system Python, any version installed via
[python.org](https://www.python.org/downloads/) or Homebrew (`brew install python`)
that is 3.10 or later will work.

### Linux

The setup works on Linux, but Playwright's Chromium browser may require
additional system libraries on minimal or server environments. If scraping
fails with a browser launch error, run the following in a terminal:

```bash
# Debian / Ubuntu
sudo apt-get install -y libglib2.0-0 libnss3 libnspr4 libdbus-1-3 \
  libatk1.0-0 libatk-bridge2.0-0 libcups2 libdrm2 libxkbcommon0 \
  libxcomposite1 libxdamage1 libxfixes3 libxrandr2 libgbm1 libasound2

# Fedora / RHEL
sudo dnf install -y nss nspr atk at-spi2-atk cups-libs libdrm \
  libxkbcommon libXcomposite libXdamage libXfixes libXrandr mesa-libgbm alsa-lib
```

Or use Playwright's built-in dependency installer. Run this once in R after
the Python environment is active:

```{r, purl=FALSE}
reticulate::py_run_string("from playwright.__main__ import main; import sys; sys.argv = ['playwright', 'install-deps', 'chromium']; main()")
```

---

## 6. Common issues and fixes

### Python is not initialized or packages appear missing

Run:

```{r, purl=FALSE}
downballot_use_python()
downballot_python_status()
```


If packages are still missing:

```{r, purl=FALSE}
downballot_install_python(reinstall = TRUE)
```


---

### “reticulate is already initialized to a different Python interpreter”

This occurs when Python is initialized **before** calling
`downballot_use_python()`.

Fix:

1. Restart your R session
2. Immediately run:

```{r, purl=FALSE}
downballot_use_python()
```

Then continue using the package.


### Reinstall everything from scratch

If the environment becomes corrupted or inconsistent:

```{r, purl=FALSE}
downballot_install_python(python = reticulate::virtualenv_starter(), reinstall = TRUE)
downballot_use_python()
downballot_python_status()
```

---

## 7. Recommended workflow summary

### First-time setup:

```{r, purl=FALSE}
reticulate::install_python()
downballot_install_python(python = reticulate::virtualenv_starter())
downballot_use_python()
```

### Every new R session:

```{r, purl=FALSE}
downballot_use_python()
```


### Check environment status:

```{r, purl=FALSE}
downballot_python_status()
```



