---
title: "Scraping election data with DownBallotR"
output: rmarkdown::html_vignette
---

```{r, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(
  purl     = FALSE,
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Prerequisites

Before scraping data you need a working Python environment. If you have not
done so yet, follow the steps in the
[Python setup vignette](https://gchickering21.github.io/DownBallotR/articles/python-setup.html):

```{r, purl=FALSE}
# One-time setup (downloads ~100-200 MB the first time)
downballot_install_python()
```

At the start of each R session:

```{r, eval=FALSE, purl=FALSE}
library(DownBallotR)
library(dplyr)

downballot_use_python()
```

---

## The single entry point: `scrape_elections()`

All data retrieval flows through one function:

```r
scrape_elections(state, ...)
```

The scraper is selected **automatically** based on `state` —
you never need to specify a backend by name:

| Priority | Condition | Scraper used |
|---|---|---|
| 1 | `state` is North Carolina | North Carolina State Board of Elections scraper (2000–present) |
| 2 | `state` is Connecticut | Connecticut CTEMS scraper (2016–present) |
| 3 | `state` is Georgia | Georgia Secretary of State scraper (2000–present) |
| 4 | `state` is Utah | Utah election results scraper (2023–present) |
| 5 | `state` is Indiana | Indiana voters portal scraper (2019–present) |
| 6 | `state` is Louisiana | Louisiana Secretary of State scraper (1982–present) |
| 7 | All other states | ElectionStats multi-state scraper |

---

## Data availability

Use `db_available_years()` to check available year ranges programmatically:

```{r, purl=FALSE}
# All sources
db_available_years()

# Filter to one state
db_available_years(state = "Virginia")
```

| Source | State | Start year | End year |
|---|---|---|---|
| ElectionStats | Vermont | 1789 | 2024 |
| ElectionStats | Virginia | 1789 | 2025 |
| ElectionStats | Colorado | 1902 | 2024 |
| ElectionStats | Massachusetts | 1970 | 2026 |
| ElectionStats | New Hampshire | 1970 | 2024 |
| ElectionStats | Idaho | 1990 | 2024 |
| ElectionStats | New York | 1994 | 2024 |
| ElectionStats | New Mexico | 2000 | 2024 |
| ElectionStats | South Carolina | 2008 | 2025 |
| North Carolina State Board of Elections | North Carolina | 2000 | present |
| Connecticut CTEMS | Connecticut | 2016 | present |
| Georgia Secretary of State | Georgia | 2000 | present |
| Utah elections site | Utah | 2023 | present |
| Indiana voters portal | Indiana | 2019 | present |
| Louisiana Secretary of State | Louisiana | 1982 | present |

To see which states are supported by the ElectionStats scraper:

```{r, purl=FALSE}
db_list_states("election_stats")
#> [1] "Colorado"        "Idaho"           "Massachusetts"   "New Hampshire"
#> [5] "New Mexico"      "New York"        "South Carolina"  "Vermont"
#> [9] "Virginia"
```

---

## Choosing the right scraper

> **Note on `level`:** The `level` argument selects the **constituency level**
> of the returned results — i.e., the geographic unit at which votes are
> tabulated. For example, `level = "state"` returns state-level constituency
> (statewide) results, `level = "county"` returns county-level constituency
> results, and `level = "precinct"` returns precinct-level constituency
> results. This is **not** the same as office level (Federal / State / Local).

| Goal | Call |
|---|---|
| Candidate + vote totals for Colorado, Massachusetts, New Hampshire, Idaho, Vermont (classic) | `scrape_elections(state = "virginia", ...)` |
| Precinct-level results for Colorado, Massachusetts, Idaho (classic) | `scrape_elections(state = "colorado", year_from = 2022, year_to = 2022, level = "precinct")` |
| Candidate + vote totals for South Carolina, New Mexico, Virginia, New York (Playwright) | `scrape_elections(state = "south_carolina", ...)` |
| Precinct-level results for South Carolina, New Mexico, Virginia (v2 CSV API) | `scrape_elections(state = "new_mexico", year_from = 2024, year_to = 2024, level = "precinct")` |
| North Carolina precinct-level local results | `scrape_elections(state = "north_carolina", year_from = 2025, year_to = 2025)` |
| Connecticut statewide + town results | `scrape_elections(state = "connecticut", year_from = 2024, year_to = 2024)` |
| Connecticut statewide totals only (faster) | `scrape_elections(state = "connecticut", year_from = 2024, year_to = 2024, level = "state")` |
| Georgia statewide + county + precinct results | `scrape_elections(state = "georgia", year_from = 2024, year_to = 2024)` |
| Georgia statewide + county only (faster) | `scrape_elections(state = "georgia", year_from = 2024, year_to = 2024, level = "county")` |
| Georgia with vote-method breakdown | `scrape_elections(state = "georgia", year_from = 2024, year_to = 2024, include_vote_methods = TRUE)` |
| Georgia precinct-level only | `scrape_elections(state = "georgia", year_from = 2024, year_to = 2024, level = "precinct")` |
| Utah statewide + county + precinct results | `scrape_elections(state = "utah", year_from = 2024, year_to = 2024)` |
| Utah statewide + county only (faster) | `scrape_elections(state = "utah", year_from = 2024, year_to = 2024, level = "county")` |
| Utah statewide only (fastest) | `scrape_elections(state = "utah", year_from = 2024, year_to = 2024, level = "state")` |
| Utah precinct-level only | `scrape_elections(state = "utah", year_from = 2024, year_to = 2024, level = "precinct")` |
| Indiana statewide + county General Election results | `scrape_elections(state = "indiana", year_from = 2024, year_to = 2024)` |
| Indiana statewide only (faster) | `scrape_elections(state = "indiana", year_from = 2022, year_to = 2022, level = "state")` |
| Louisiana statewide + parish results | `scrape_elections(state = "louisiana", year_from = 2024, year_to = 2024)` |
| Louisiana statewide only (faster; skips parish scraping) | `scrape_elections(state = "louisiana", year_from = 2023, year_to = 2023, level = "state")` |

---

## Next: analyzing the data

Once you have a scraped data frame, the
[Example analyses vignette](https://gchickering21.github.io/DownBallotR/articles/example-analyses.html)
walks through worked example questions — wins by party, closest races,
down-ballot drop-off, county-level variation, and cross-state
comparisons — with tidyverse code you can adapt to your own scrapes.

---

## Detailed articles

Each scraper has its own article with full argument documentation,
worked examples, and column descriptions (available on the package website):

- [ElectionStats states](https://gchickering21.github.io/DownBallotR/articles/election-stats.html) —
  Virginia, Massachusetts, Colorado, New Hampshire, Idaho, Vermont (classic) +
  South Carolina, New Mexico, New York (Playwright); candidate + county + precinct results
- [North Carolina](https://gchickering21.github.io/DownBallotR/articles/north-carolina.html) —
  North Carolina State Board of Elections; precinct-level results
- [Connecticut](https://gchickering21.github.io/DownBallotR/articles/connecticut.html) —
  Connecticut CTEMS; statewide + town results, 2016–present
- [Georgia](https://gchickering21.github.io/DownBallotR/articles/georgia.html) —
  Georgia Secretary of State; statewide + county + precinct results, 2000–present
- [Utah](https://gchickering21.github.io/DownBallotR/articles/utah.html) —
  Utah elections site; statewide + county + precinct results, 2023–present
- [Indiana](https://gchickering21.github.io/DownBallotR/articles/indiana.html) —
  Indiana voters portal; statewide + county General Election results, 2019–present
- [Louisiana](https://gchickering21.github.io/DownBallotR/articles/louisiana.html) —
  Louisiana Secretary of State; statewide + parish results, 1982–present

---

## Performance tips

- **Start small** — test with a single year and state before requesting
  large date ranges.
- **Playwright states are slower** — South Carolina, New Mexico, and New York
  launch a headless browser for each scrape. Expect several seconds per year.
- **Parallel scraping is on by default** for classic (requests-based)
  ElectionStats states. Pass `parallel = FALSE` to disable.
- **Be polite** — built-in delays are intentional. Do not reduce or
  disable them; excessive requests may result in temporary IP blocks.
